### 2020-01-18 Temp_tables/TablesForGoogleDoc.tex
Initial Overleaf Import

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\end{document}