### 2020-01-18 Temp_tables/table_3.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:55:48
\begin{table}[!htbp] \centering 
  \caption{Additional controls} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
\cline{2-5} 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.275$^{*}$ & $-$0.320$^{**}$ & $-$0.331$^{**}$ & $-$0.378$^{**}$ \\ 
  & (0.166) & (0.145) & (0.155) & (0.160) \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i  & 0.081 & 0.070 & 0.004 & 0.019 \\ 
  & (0.158) & (0.147) & (0.155) & (0.161) \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i  & $-$0.083 & $-$0.034 & $-$0.034 & $-$0.044 \\ 
  & (0.148) & (0.150) & (0.153) & (0.161) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{count share SOE}_{i}  & 1.531 &  &  &  \\ 
  & (1.185) &  &  &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{count share SOE}_{i}  & 0.181 &  &  &  \\ 
  & (1.054) &  &  &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{count share SOE}_{i}  & 0.077 &  &  &  \\ 
  & (1.091) &  &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  &  & 1.305$^{*}$ &  &  \\ 
  &  & (0.749) &  &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  &  & 0.300 &  &  \\ 
  &  & (0.689) &  &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  &  & $-$0.327 &  &  \\ 
  &  & (0.774) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  &  & 0.924 &  \\ 
  &  &  & (0.595) &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  &  & 0.477 &  \\ 
  &  &  & (0.544) &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  &  & $-$0.150 &  \\ 
  &  &  & (0.559) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  &  & 1.391$^{*}$ \\ 
  &  &  &  & (0.725) \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  &  & 0.525 \\ 
  &  &  &  & (0.666) \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  &  & $-$0.142 \\ 
  &  &  &  & (0.723) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.851 & 0.851 & 0.851 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{4}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}