### 2020-01-18 Temp_tables/table_4.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:55:57
\begin{table}[!htbp] \centering 
  \caption{Additional controls (2)} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
\cline{2-5} 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
   \text{Polluted}_i \times \text{ln gdp per cap)}_{ct}  & $-$0.038 & $-$0.034 & $-$0.029 & $-$0.036 \\ 
  & (0.297) & (0.298) & (0.298) & (0.297) \\ 
   \text{Polluted}_i \times \text{ln population)}_{ct}  & 0.165 & 0.170 & 0.172 & 0.172 \\ 
  & (0.227) & (0.228) & (0.228) & (0.228) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.238 & $-$0.280$^{**}$ & $-$0.317$^{**}$ & $-$0.360$^{**}$ \\ 
  & (0.153) & (0.142) & (0.151) & (0.156) \\ 
   TCZ_c \times \text{Period} \times \text{count share SOE}_{i}  & $-$1.177 &  &  &  \\ 
  & (0.888) &  &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{count share SOE}_{i}  & 1.522 &  &  &  \\ 
  & (1.115) &  &  &  \\ 
   TCZ_c \times \text{Period} \times \text{output share SOE}_{i}  &  & $-$0.842 &  &  \\ 
  &  & (0.574) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  &  & 1.327$^{*}$ &  &  \\ 
  &  & (0.731) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share SOE}_{i}  &  &  & $-$0.707 &  \\ 
  &  &  & (0.438) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  &  & 1.074$^{*}$ &  \\ 
  &  &  & (0.577) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share SOE}_{i}  &  &  &  & $-$1.044$^{*}$ \\ 
  &  &  &  & (0.562) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  &  & 1.545$^{**}$ \\ 
  &  &  &  & (0.709) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes \\ 
Observations & 30,195 & 30,195 & 30,195 & 30,195 \\ 
R$^{2}$ & 0.852 & 0.852 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{4}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}