### 2020-01-18 Temp_tables/table_5.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:56:07
\begin{table}[!htbp] \centering 
  \caption{SOE dominated sectors} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
\cline{2-5} 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Polluted}_i \times \text{Period}  & $-$0.123 & 0.101 & 0.114 & 0.235 \\ 
  & (0.216) & (0.234) & (0.229) & (0.232) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes \\ 
Observations & 9,232 & 9,824 & 10,423 & 10,076 \\ 
R$^{2}$ & 0.888 & 0.865 & 0.869 & 0.862 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{4}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}