### 2020-01-18 Temp_tables/table_6.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[htbp]\centering
    \resizebox{1\textwidth}{!}{
        \begin{threeparttable}  
            \caption{\small Summary Statistics by City-industry characteristics and City characteristics}
            \begin{tabular}{lrrrr}
                \multicolumn{1}{l}{\textbf{Panel A: SOE dominated sector}} \\
                    \toprule
                    & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
                    \cline{2-5} 
                    \\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
                    \\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\
                    \hline \\[-1.8ex] 
                    TCZ_c \times \text{Polluted}_i \times \text{Period}  & $-$0.123 & 0.101 & 0.114 & 0.235 \\ 
                    & (0.216) & (0.234) & (0.229) & (0.232) \\ 
                    \hline \\[-1.8ex] 
                    City-year fixed effects & Yes & Yes & Yes & Yes \\ 
                    Industry-year fixed effects & Yes & Yes & Yes & Yes \\
                    City-industry fixed effects & Yes & Yes & Yes & Yes \\
                    Observations & 9,232 & 9,824 & 10,423 & 10,076 \\ 
                    Number sectors & X & 92 & 104& 98 \\
                    R$^{2}$ & 0.888 & 0.865 & 0.869 & 0.862 \\
                    \bottomrule
                    \\ %%% Create second table
                \multicolumn{1}{l}{\textbf{Panel B: SOE dominating sector}} \\
                    \toprule
                    & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
                    \cline{2-5} 
                    \\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
                    \\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\
                    \hline \\[-1.8ex] 
                    TCZ_c \times \text{Polluted}_i \times \text{Period}  & $-$0.138 & $-$0.243$^{**}$ & $-$0.272$^{**}$ & $-$0.284$^{**}$ \\ 
                    & (0.145) & (0.124) & (0.124) & (0.125) \\ 
                    \hline \\[-1.8ex] 
                    City-year fixed effects & Yes & Yes & Yes & Yes \\ 
                    Industry-year fixed effects & Yes & Yes & Yes & Yes \\
                    City-industry fixed effects & Yes & Yes & Yes & Yes \\
                    Observations & 21,444 & 20,852 & 20,253 & 20,600 \\ 
                    Number sectors & X & 204 & 192& 198 \\
                    R$^{2}$ & 0.849 & 0.857 & 0.853 & 0.858 \\ 
            \end{tabular}
            \begin{tablenotes}
                \small
                \item Panel A 
                Panel B \\
                Sources: Author's own computation 
            \end{tablenotes}
        \end{threeparttable}
    }
\end{table}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:56:15
\begin{table}[!htbp] \centering 
  \caption{SOE dominating sectors} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
\cline{2-5} 
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Polluted}_i \times \text{Period}  & $-$0.138 & $-$0.243$^{**}$ & $-$0.272$^{**}$ & $-$0.284$^{**}$ \\ 
  & (0.145) & (0.124) & (0.124) & (0.125) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes \\ 
Observations & 21,444 & 20,852 & 20,253 & 20,600 \\ 
R$^{2}$ & 0.849 & 0.857 & 0.853 & 0.858 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{4}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{4}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{4}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}