### 2020-01-18 Temp_tables/table_7.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}
% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:56:21
\begin{table}[!htbp] \centering 
  \caption{11th FYP city pollution reduction mandate} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
  \label{} 
\begin{tabular}{@{\extracolsep{5pt}}lcc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{2}{c}{The Dependent variable: Ln SO2_{cit}} \\ 
\cline{2-3} 
\\[-1.8ex] & (1) & (2)\\ 
\\[-1.8ex] & (TCZ) & (No TCZ)\\
\hline \\[-1.8ex] 
   \text{Period} \times target_c \times \text{Polluted}_i  & $-$0.413$^{***}$ & $-$1.415 \\ 
  & (0.141) & (1.672) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes \\ 
Observations & 23,333 & 7,343 \\ 
R$^{2}$ & 0.847 & 0.892 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{2}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{2}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{2}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}