### 2020-01-19 Temp_tables/table_4.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[!htbp] \centering
  \caption{Additional controls (2)}
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
    \label{}
    \begin{tabular}{@{\extracolsep{5pt}}lcccc}
      \\[-1.8ex]\hline
      \hline \\[-1.8ex]
      & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{ikt}} \\
      \cline{2-5}
      \\[-1.8ex] & (1) & (2) & (3) & (4)\\
      \hline \\[-1.8ex]
      \text{Polluted}_k \times \text{ln gdp per cap}_{it}                                    & $-$0.038 & $-$0.034        & $-$0.029        & $-$0.036        \\
                                                                                              & (0.297)  & (0.298)         & (0.298)         & (0.297)         \\
      \text{Polluted}_k \times \text{ln population}_{it}                                     & 0.165    & 0.170           & 0.172           & 0.172           \\
                                                                                              & (0.227)  & (0.228)         & (0.228)         & (0.228)         \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k                                     & $-$0.238 & $-$0.280$^{**}$ & $-$0.317$^{**}$ & $-$0.360$^{**}$ \\
                                                                                              & (0.153)  & (0.142)         & (0.151)         & (0.156)         \\
      TCZ_i \times \text{Period} \times \text{count share SOE}_{k}                            & $-$1.177 &                 &                 &                 \\
                                                                                              & (0.888)  &                 &                 &                 \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{count share SOE}_{k}   & 1.522    &                 &                 &                 \\
                                                                                              & (1.115)  &                 &                 &                 \\
      TCZ_i \times \text{Period} \times \text{output share SOE}_{k}                           &          & $-$0.842        &                 &                 \\
                                                                                              &          & (0.574)         &                 &                 \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{output share SOE}_{k}  &          & 1.327$^{*}$     &                 &                 \\
                                                                                              &          & (0.731)         &                 &                 \\
      TCZ_i \times \text{Period} \times \text{capital share SOE}_{k}                          &          &                 & $-$0.707        &                 \\
                                                                                              &          &                 & (0.438)         &                 \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{capital share SOE}_{k} &          &                 & 1.074$^{*}$     &                 \\
                                                                                              &          &                 & (0.577)         &                 \\
      TCZ_i \times \text{Period} \times \text{labour share SOE}_{k}                           &          &                 &                 & $-$1.044$^{*}$  \\
                                                                                              &          &                 &                 & (0.562)         \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{labour share SOE}_{k}  &          &                 &                 & 1.545$^{**}$    \\
                                                                                              &          &                 &                 & (0.709)         \\
      \hline \\[-1.8ex]
      City-year fixed effects                                                                 & Yes      & Yes             & Yes             & Yes             \\
      Industry-year fixed effects                                                             & Yes      & Yes             & Yes             & Yes             \\
      City-industry fixed effects                                                             & Yes      & Yes             & Yes             & Yes             \\
      Observations                                                                            & 30,195   & 30,195          & 30,195          & 30,195          \\
      R$^{2}$                                                                                 & 0.852    & 0.852           & 0.852           & 0.852           \\
      \hline
      \hline \\[-1.8ex]
      \end{tabular}
  \end{adjustbox}
  \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
    \end{tablenotes}
\end{table}

\end{document}