### 2020-01-19 Temp_tables/table_6.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[!htbp] \centering
  %\resizebox{1\textwidth}{!}{
    %\begin{threeparttable}
    \caption{SOE}
      \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
     \label{}
      \begin{tabular}{lrrrr}
        \multicolumn{1}{l}{\textbf{Panel A: SOE Dominating sector}} \\
        \toprule
        & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{ikt}} \\
        \cline{2-5}
        \\[-1.8ex] & (1) & (2) & (3) & (4)\\
        \\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\
        \hline \\[-1.8ex]
        TCZ_i \times \text{Polluted}_k \times \text{Period} & $-$0.123 & 0.101           & 0.114           & 0.235           \\
                                                            & (0.216)  & (0.234)         & (0.229)         & (0.232)         \\
        \hline \\[-1.8ex]
        City-year fixed effects                             & Yes      & Yes             & Yes             & Yes             \\
        Industry-year fixed effects                         & Yes      & Yes             & Yes             & Yes             \\
        City-industry fixed effects                         & Yes      & Yes             & Yes             & Yes             \\
        Observations                                        & 9,232    & 9,824           & 10,423          & 10,076          \\
        Number sectors                                      & X        & 92              & 104             & 98              \\
        R$^{2}$                                             & 0.888    & 0.865           & 0.869           & 0.862           \\
        \bottomrule
        \\ %%% Create second table
        \multicolumn{1}{l}{\textbf{Panel B: Prviate Dominating sectors}} \\
        \toprule
        & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{ikt}} \\
        \cline{2-5}
        \\[-1.8ex] & (1) & (2) & (3) & (4)\\
        \\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\
        \hline \\[-1.8ex]
        TCZ_i \times \text{Polluted}_k \times \text{Period} & $-$0.138 & $-$0.243$^{**}$ & $-$0.272$^{**}$ & $-$0.284$^{**}$ \\
                                                            & (0.145)  & (0.124)         & (0.124)         & (0.125)         \\
        \hline \\[-1.8ex]
        City-year fixed effects                             & Yes      & Yes             & Yes             & Yes             \\
        Industry-year fixed effects                         & Yes      & Yes             & Yes             & Yes             \\
        City-industry fixed effects                         & Yes      & Yes             & Yes             & Yes             \\
        Observations                                        & 21,444   & 20,852          & 20,253          & 20,600          \\
        Number sectors                                      & X        & 204             & 192             & 198             \\
        R$^{2}$                                             & 0.849    & 0.857           & 0.853           & 0.858           \\
    \end{tabular}
    \end{adjustbox}
    \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
      \\
    \end{tablenotes}
\end{table}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:56:15
\begin{table}[!htbp] \centering
  \caption{SOE dominating sectors}
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
    \label{}
    \begin{tabular}{@{\extracolsep{5pt}}lcccc}
      \\[-1.8ex]\hline
      \hline \\[-1.8ex]
      & \multicolumn{4}{c}{The Dependent variable: Ln SO2_{ikt}} \\
      \cline{2-5}
      \\[-1.8ex] & (1) & (2) & (3) & (4)\\
      \\[-1.8ex] & (SOE count) & (Output share) & (Capital share) & (Labour share)\\
      \hline \\[-1.8ex]
      TCZ_i \times \text{Polluted}_k \times \text{Period} & $-$0.138 & $-$0.243$^{**}$ & $-$0.272$^{**}$ & $-$0.284$^{**}$ \\
                                                          & (0.145)  & (0.124)         & (0.124)         & (0.125)         \\
      \hline \\[-1.8ex]
      City-year fixed effects                             & Yes      & Yes             & Yes             & Yes             \\
      Industry-year fixed effects                         & Yes      & Yes             & Yes             & Yes             \\
      City-industry fixed effects                         & Yes      & Yes             & Yes             & Yes             \\
      Observations                                        & 21,444   & 20,852          & 20,253          & 20,600          \\
      R$^{2}$                                             & 0.849    & 0.857           & 0.853           & 0.858           \\
      \hline
      \hline \\[-1.8ex]
      \end{tabular}
  \end{adjustbox}
  \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
    \end{tablenotes}
\end{table}

\end{document}