### 2020-01-19 Temp_tables/table_7.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[!htbp] \centering
  \caption{11th FYP city pollution reduction mandate}
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
    \label{}
    \begin{tabular}{@{\extracolsep{5pt}}lcc}
      \\[-1.8ex]\hline
      \hline \\[-1.8ex]
      & \multicolumn{2}{c}{The Dependent variable: Ln SO2_{ikt}} \\
      \cline{2-3}
      \\[-1.8ex] & (1) & (2)\\
      \\[-1.8ex] & (TCZ) & (No TCZ)\\
      \hline \\[-1.8ex]
      \text{Period} \times target_i \times \text{Polluted}_k & $-$0.413$^{***}$ & $-$1.415 \\
                                                             & (0.141)          & (1.672)  \\
      \hline \\[-1.8ex]
      City-year fixed effects                                & Yes              & Yes      \\
      Industry-year fixed effects                            & Yes              & Yes      \\
      City-industry fixed effects                            & Yes              & Yes      \\
      Observations                                           & 23,333           & 7,343    \\
      R$^{2}$                                                & 0.847            & 0.892    \\
      \hline
      \hline \\[-1.8ex]
      \end{tabular}
  \end{adjustbox}
  \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
    \end{tablenotes}
\end{table}

\end{document}