### 2020-04-27 Temp_tables/Tables_paper/01_Paper_version_1/table_3.tex
back up before new paper

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Dec 08, 2019 - 08:55:48
\begin{table}[!htbp] \centering
  \caption{Soft budget constraint as reflected in the SOEs’ reaction to the SPZs’ and coastal area’s policies in China}
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
    \label{tab:table5}
    \begin{tabular}{@{\extracolsep{5pt}}lcccc}
      \\[-1.8ex]\hline
      \hline \\[-1.8ex]
      & \multicolumn{3}{c}{The Dependent variable: Ln SO2_{ikt}} \\
      \cline{2-4}
      \\[-1.8ex] & Output & Capital & Labour \\
      \\[-1.8ex] & (1) & (2) & (3) \\
      \hline \\[-1.8ex]
      TCZ_i \times \text{Period} \times \text{Polluted}_k                                         & $-$0.320$^{**}$ & $-$0.331$^{**}$ & $-$0.378$^{**}$ \\
                                                                                                  & (0.145)         & (0.155)         & (0.160)         \\
      SPZ_i \times \text{Period} \times \text{Polluted}_k                                         & 0.070           & 0.004           & 0.019           \\
                                                                                                  & (0.147)         & (0.155)         & (0.161)         \\
      Coastal_i \times \text{Period} \times \text{Polluted}_k                                     & $-$0.034        & $-$0.034        & $-$0.044        \\
                                                                                                  & (0.150)         & (0.153)         & (0.161)         \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{output share SOE}_{k}      & 1.305$^{*}$     &                 &                 \\
                                                                                                  & (0.749)         &                 &                 \\
      SPZ_i \times \text{Period} \times \text{Polluted}_k \times \text{output share SOE}_{k}      & 0.300           &                 &                 \\
                                                                                                  & (0.689)         &                 &                 \\
      Coastal_i \times \text{Period} \times \text{Polluted}_k \times \text{output share SOE}_{k}  & $-$0.327        &                 &                 \\
                                                                                                  & (0.774)         &                 &                 \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{capital share SOE}_{k}     &                 & 0.924           &                 \\
                                                                                                  &                 & (0.595)         &                 \\
      SPZ_i \times \text{Period} \times \text{Polluted}_k \times \text{capital share SOE}_{k}     &                 & 0.477           &                 \\
                                                                                                  &                 & (0.544)         &                 \\
      Coastal_i \times \text{Period} \times \text{Polluted}_k \times \text{capital share SOE}_{k} &                 & $-$0.150        &                 \\
                                                                                                  &                 & (0.559)         &                 \\
      TCZ_i \times \text{Period} \times \text{Polluted}_k \times \text{labour share SOE}_{k}      &                 &                 & 1.391$^{*}$     \\
                                                                                                  &                 &                 & (0.725)         \\
      SPZ_i \times \text{Period} \times \text{Polluted}_k \times \text{labour share SOE}_{k}      &                 &                 & 0.525           \\
                                                                                                  &                 &                 & (0.666)         \\
      Coastal_i \times \text{Period} \times \text{Polluted}_k \times \text{labour share SOE}_{k}  &                 &                 & $-$0.142        \\
                                                                                                  &                 &                 & (0.723)         \\
      \hline \\[-1.8ex]
      City-year fixed effects                                                                     & Yes             & Yes             & Yes             \\
      Industry-year fixed effects                                                                 & Yes             & Yes             & Yes             \\
      City-industry fixed effects                                                                 & Yes             & Yes             & Yes             \\
      Observations                                                                                & 30,676          & 30,676          & 30,676          \\
      R$^{2}$                                                                                     & 0.851           & 0.851           & 0.852           \\
      \hline
      \hline \\[-1.8ex]
      \end{tabular}
  \end{adjustbox}
  \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
    \end{tablenotes}
\end{table}

\end{document}