### 2020-04-27 Temp_tables/Tables_paper/01_Paper_version_1/table_8.tex

\documentclass[12pt]{article}

\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[!htbp] \centering
  \caption{Fighting against pollution via productivity improvement: SOEs versus private firms}
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
    \label{eq:tfp}
    \begin{tabular}{@{\extracolsep{5pt}}lcccccccc}
      \\[-1.8ex]\hline
      \hline \\[-1.8ex]
      & \multicolumn{5}{c}{The Dependent variable:} \\
      \\[-1.8ex] & (SOE) & (PRIVATE) & (SOE) & (PRIVATE) & (SOE) & (PRIVATE) & (SOE) & (PRIVATE)\\
      \\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\
      \cline{2-8}
      TCZ_c \times \text{Polluted}_i \times \text{Period}     & 0.506$^{**}$ & 0.536$^{***}$ &               &               &               &               & 0.005         & 0.345$^{***}$ \\
                                                              & (0.205)      & (0.023)       &               &               &               &               & (0.213)       & (0.111)       \\
      SPZ_c \times \text{Polluted}_i \times \text{Period}     &              &               & 0.070$^{***}$ & 0.147$^{***}$ &               &               & 0.070$^{***}$ & 0.147$^{***}$ \\
                                                              &              &               & (0.016)       & (0.001)       &               &               & (0.016)       & (0.001)       \\
      Coastal_c \times \text{Polluted}_i \times \text{Period} &              &               &               &               & 0.513$^{***}$ & 0.382$^{***}$ & 0.510$^{***}$ & 0.191$^{*}$   \\
                                                              &              &               &               &               & (0.159)       & (0.116)       & (0.171)       & (0.112)       \\
      \hline \\[-1.8ex]
      City-year fixed effects                                 & Yes          & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           \\
      Industry-year fixed effects                             & Yes          & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           \\
      City-industry fixed effects                             & Yes          & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           & Yes           \\
      Observations                                            & 2,415        & 34,437        & 2,415         & 34,437        & 2,415         & 34,437        & 2,415         & 34,437        \\
      R$^{2}$                                                 & 0.482        & 0.104         & 0.481         & 0.103         & 0.482         & 0.104         & 0.482         & 0.104         \\
      \hline
      \hline \\[-1.8ex]
      \end{tabular}
  \end{adjustbox}
  \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
    \end{tablenotes}
\end{table}

\end{document}