### 2020-04-27 Temp_tables/Tables_paper/02_paper_version_2/00_baseline_SOE.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 26, 2020 - 09:48:44
\begin{table}[!htbp] \centering 
  \caption{Table Baseline Panel 1 - SOE} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-9}

\\[-1.8ex]
            &\multicolumn{2}{c}{Full sample}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{Employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
  output_{cit} & 0.144 & $-$0.066 & 0.875$^{*}$ & 0.216 & 1.092$^{**}$ & 0.308 & 0.995$^{**}$ & 0.087 \\ 
  & (0.152) & (0.089) & (0.457) & (0.257) & (0.464) & (0.397) & (0.421) & (0.280) \\ 
  capital_{cit} & 1.481$^{**}$ & 0.971$^{***}$ & $-$3.614$^{***}$ & $-$0.611 & $-$4.106$^{***}$ & $-$0.485 & $-$4.094$^{***}$ & $-$0.393 \\ 
  & (0.730) & (0.366) & (1.088) & (0.617) & (1.146) & (0.773) & (1.104) & (0.708) \\ 
  labour_{cit} & 3.089$^{***}$ & 1.538$^{**}$ & 11.671$^{***}$ & 4.601$^{***}$ & 12.228$^{***}$ & 5.216$^{***}$ & 12.327$^{***}$ & 4.783$^{***}$ \\ 
  & (0.832) & (0.767) & (2.326) & (1.475) & (2.374) & (1.353) & (2.359) & (1.548) \\ 
   target_c \times \text{Period}  & $-$0.003 &  & 0.087 &  & 0.012 &  & 0.128 &  \\ 
  & (0.104) &   & (0.381) &   & (0.379) &   & (0.379) &   \\ 
   target_c \times \text{Polluted}_i  & 0.438$^{***}$ &  & 0.838$^{**}$ &  & 0.748$^{*}$ &  & 0.839$^{**}$ &  \\ 
  & (0.140) &   & (0.383) &   & (0.390) &   & (0.379) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.291$^{**}$ & $-$0.430$^{***}$ & $-$0.430 & 0.009 & $-$0.344 & $-$0.159 & $-$0.468 & $-$0.215 \\ 
  & (0.146) & (0.132) & (0.465) & (0.465) & (0.443) & (0.435) & (0.432) & (0.461) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Observations & 30,676 & 30,676 & 9,165 & 9,165 & 9,149 & 9,149 & 9,011 & 9,011 \\ 
R$^{2}$ & 0.346 & 0.851 & 0.377 & 0.872 & 0.376 & 0.868 & 0.373 & 0.869 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1, 3, 5, 7).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}