### 2020-04-27 Temp_tables/Tables_paper/02_paper_version_2/01_baseline__No_SOE.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 26, 2020 - 09:48:49
\begin{table}[!htbp] \centering 
  \caption{Table Baseline Panel 2 - No SOE} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-9}

\\[-1.8ex]
            &\multicolumn{2}{c}{Full sample}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{Employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
  output_{cit} & 0.144 & $-$0.066 & 0.116 & $-$0.077 & 0.107 & $-$0.077 & 0.113 & $-$0.073 \\ 
  & (0.152) & (0.089) & (0.134) & (0.094) & (0.130) & (0.090) & (0.130) & (0.095) \\ 
  capital_{cit} & 1.481$^{**}$ & 0.971$^{***}$ & 1.885$^{***}$ & 0.942$^{**}$ & 1.930$^{***}$ & 0.905$^{**}$ & 1.744$^{***}$ & 0.911$^{**}$ \\ 
  & (0.730) & (0.366) & (0.575) & (0.445) & (0.557) & (0.401) & (0.551) & (0.441) \\ 
  labour_{cit} & 3.089$^{***}$ & 1.538$^{**}$ & 2.592$^{***}$ & 1.423$^{*}$ & 2.567$^{***}$ & 1.329$^{*}$ & 2.611$^{***}$ & 1.376 \\ 
  & (0.832) & (0.767) & (0.718) & (0.798) & (0.707) & (0.778) & (0.731) & (0.847) \\ 
   target_c \times \text{Period}  & $-$0.003 &  & $-$0.050 &  & $-$0.031 &  & $-$0.049 &  \\ 
  & (0.104) &   & (0.105) &   & (0.105) &   & (0.105) &   \\ 
   target_c \times \text{Polluted}_i  & 0.438$^{***}$ &  & 0.388$^{***}$ &  & 0.394$^{***}$ &  & 0.369$^{**}$ &  \\ 
  & (0.140) &   & (0.148) &   & (0.145) &   & (0.150) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.291$^{**}$ & $-$0.430$^{***}$ & $-$0.250$^{*}$ & $-$0.429$^{***}$ & $-$0.284$^{*}$ & $-$0.430$^{***}$ & $-$0.256$^{*}$ & $-$0.434$^{***}$ \\ 
  & (0.146) & (0.132) & (0.150) & (0.141) & (0.151) & (0.140) & (0.151) & (0.144) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Observations & 30,676 & 30,676 & 21,511 & 21,511 & 21,527 & 21,527 & 21,665 & 21,665 \\ 
R$^{2}$ & 0.346 & 0.851 & 0.355 & 0.855 & 0.359 & 0.856 & 0.358 & 0.856 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1, 3, 5, 7).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}