### 2020-04-27 Temp_tables/Tables_paper/02_paper_version_2/04_DC_TCZ_C_SPZ.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 26, 2020 - 10:56:13
\begin{table}[!htbp] \centering 
  \caption{\\ Effects of the policies: TCZ, Go West, and SPZ } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{12}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-13}

\\[-1.8ex]
            &\multicolumn{2}{c}{TCZ}&\multicolumn{2}{c}{No TCZ}&\multicolumn{2}{c}{ Coastal }&\multicolumn{2}{c}{No  Coastal}&\multicolumn{2}{c}{ SPZ }&\multicolumn{2}{c}{No  SPZ}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12)\\ 
\hline \\[-1.8ex] 
   target_c \times \text{Period}  & $-$0.012 &  & 0.899 &  & $-$0.202$^{*}$ &  & 0.470$^{**}$ &  & $-$0.031 &  & 0.731$^{***}$ &  \\ 
  & (0.111) &   & (0.874) &   & (0.111) &   & (0.224) &   & (0.114) &   & (0.225) &   \\ 
   target_c \times \text{Polluted}_i  & 0.388$^{***}$ &  & 2.570$^{**}$ &  & 0.321$^{**}$ &  & 0.661$^{***}$ &  & 0.446$^{***}$ &  & 0.688$^{**}$ &  \\ 
  & (0.146) &   & (1.182) &   & (0.161) &   & (0.238) &   & (0.158) &   & (0.272) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.254 & $-$0.413$^{***}$ & $-$1.335 & $-$1.415 & $-$0.241 & $-$0.356$^{**}$ & $-$0.424 & $-$0.714$^{**}$ & $-$0.317$^{*}$ & $-$0.412$^{***}$ & $-$0.695$^{**}$ & $-$0.844$^{***}$ \\ 
  & (0.155) & (0.138) & (1.128) & (1.059) & (0.174) & (0.157) & (0.274) & (0.278) & (0.177) & (0.157) & (0.302) & (0.324) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Observations & 23,333 & 23,333 & 7,343 & 7,343 & 17,915 & 17,915 & 12,761 & 12,761 & 14,513 & 14,513 & 16,163 & 16,163 \\ 
R$^{2}$ & 0.346 & 0.847 & 0.416 & 0.892 & 0.359 & 0.852 & 0.373 & 0.870 & 0.340 & 0.846 & 0.377 & 0.871 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1, 3, 5, 7, 9, 11).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}