### 2020-04-27 Temp_tables/Tables_paper/02_paper_version_2/05_DC_concentrated.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 26, 2020 - 11:08:20
\begin{table}[!htbp] \centering 
  \caption{\\ Industrial concentration and environmental regulation effectiveness} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-5}

\\[-1.8ex]
            &\multicolumn{2}{c}{Concentrated}&\multicolumn{2}{c}{No Concentrated}\\
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
   target_c \times \text{Period}  & $-$0.156 &  & 0.012 &  \\ 
  & (0.323) &   & (0.104) &   \\ 
   target_c \times \text{Polluted}_i  & 0.532 &  & 0.411$^{***}$ &  \\ 
  & (0.351) &   & (0.145) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.074 & 0.208 & $-$0.302$^{**}$ & $-$0.504$^{***}$ \\ 
  & (0.392) & (0.396) & (0.153) & (0.143) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes \\ 
Observations & 9,181 & 9,181 & 21,495 & 21,495 \\ 
R$^{2}$ & 0.406 & 0.888 & 0.342 & 0.846 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1 and 3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}