### 2020-04-27 Temp_tables/Tables_paper/02_paper_version_2/06_DC_kuznet.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 26, 2020 - 11:12:08
\begin{table}[!htbp] \centering 
  \caption{\\ Turning points and Kuznets curves} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-11}

\\[-1.8ex]
            &\multicolumn{2}{c}{City}&\multicolumn{2}{c}{Concentration}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{Employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& TCZ & No TCZ & Concentrated & No Concentrated & SOE dominated & SOE No dominated & SOE dominated & SOE No dominated & SOE dominated & SOE No dominated\\
 \hline \\[-1.8ex] 
  \text{(ln gdp per cap)}_{ct}  & 2.708$^{***}$ & 0.269 & $-$0.626 & 3.391$^{***}$ & $-$0.096 & 2.611$^{***}$ & 1.437 & 2.559$^{***}$ & 0.434 & 2.984$^{***}$ \\ 
  & (0.927) & (1.428) & (1.192) & (0.971) & (1.784) & (0.837) & (1.882) & (0.847) & (1.555) & (0.854) \\ 
   \text{(ln gdp per cap)}^2_{ct}  & $-$0.132$^{***}$ & 0.002 & 0.039 & $-$0.158$^{***}$ & 0.017 & $-$0.127$^{***}$ & $-$0.062 & $-$0.126$^{***}$ & $-$0.015 & $-$0.143$^{***}$ \\ 
  & (0.045) & (0.071) & (0.061) & (0.047) & (0.090) & (0.041) & (0.096) & (0.041) & (0.080) & (0.042) \\ 
   \text{(ln population)}_{ct}  & 0.062 & 0.011 & 0.328$^{*}$ & 0.026 & 0.043 & 0.040 & $-$0.093 & 0.055 & 0.054 & 0.036 \\ 
  & (0.103) & (0.168) & (0.191) & (0.102) & (0.211) & (0.096) & (0.221) & (0.094) & (0.218) & (0.092) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 28795 & - & - & 45396 & - & 30264 & - & 24867 & - & 35190 \\ 
turning point Dollar & 3568 & - & - & 5625 & - & 3750 & - & 3081 & - & 4361 \\ 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 22,865 & 7,330 & 9,153 & 21,042 & 9,160 & 21,035 & 9,137 & 21,058 & 8,993 & 21,202 \\ 
R$^{2}$ & 0.344 & 0.415 & 0.405 & 0.340 & 0.376 & 0.353 & 0.375 & 0.357 & 0.372 & 0.357 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}