### 2020-04-27 Temp_tables/Tables_paper/02_paper_version_2/XX_tfp_concetration.tex

\documentclass[preview]{standalone} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Fri, Apr 24, 2020 - 17:00:56
\begin{table}[!htbp] \centering 
  \caption{Reduction mandate and TFP : industrial concentration} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{Dependent variable \text { TFP }_{fikt}} \\ 
\cline{2-5}

\\[-1.8ex]
            &\multicolumn{2}{c}{SOE}&\multicolumn{2}{c}{PRIVATE}\\
\\[-1.8ex] & (1) & (2) & (3) & (4)\\
 \\[-1.8ex]& Concentrated & NO Concentrated & Concentrated & NO Concentrated\\
 \hline \\[-1.8ex] 
   target_c \times \text{Period} \times \text{Polluted}_i  & 0.068 & 0.159$^{**}$ & $-$0.035 & $-$0.015 \\ 
  & (0.084) & (0.063) & (0.032) & (0.024) \\ 
 \hline \\[-1.8ex] 
Firm & Yes & Yes & Yes & Yes \\ 
City-industry & No & No & Yes & Yes \\ 
City-time & Yes & Yes & Yes & Yes \\ 
time-industry & No & No & Yes & Yes \\ 
Observations & 23,054 & 18,434 & 170,305 & 437,004 \\ 
R$^{2}$ & 0.957 & 0.953 & 0.869 & 0.859 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \\
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}
\end{tablenotes}
\end{table}

\end{document}