### 2020-04-27 Temp_tables/Temp_tables_revision/00_Archive/03_foreign.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Mar 08, 2020 - 11:14:33
\begin{table}[!htbp] \centering 
  \caption{Baseline results revision: Foreign} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{The Dependent variable:} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.083 & $-$0.170 & $-$0.042 & $-$0.021 & $-$0.049 & 0.011 \\ 
  & (0.185) & (0.177) & (0.166) & (0.214) & (0.210) & (0.195) \\ 
   TCZ_c \times \text{Period} \times \text{output share Foreign}_{i} & $-$0.483 &  &  & 0.856 &  &  \\ 
  & (0.808) &  &  & (1.031) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share Foreign}_{i} & $-$0.595 &  &  & $-$0.874 &  &  \\ 
  & (1.398) &  &  & (1.539) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share Foreign}_{i}  &  & $-$0.535 &  &  & 0.650 &  \\ 
  &  & (0.781) &  &  & (0.923) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share Foreign}_{i}  &  & 0.233 &  &  & $-$0.555 &  \\ 
  &  & (1.148) &  &  & (1.296) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share Foreign}_{i}  &  &  & 0.079 &  &  & 1.437 \\ 
  &  &  & (0.750) &  &  & (0.929) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share Foreign}_{i}  &  &  & $-$0.983 &  &  & $-$1.061 \\ 
  &  &  & (1.523) &  &  & (1.572) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.346 & 0.346 & 0.346 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{6}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{6}{l}{are clustered by city} \\ 
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}