### 2020-04-27 Temp_tables/Temp_tables_revision/00_Archive/table_1_bis.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Feb 23, 2020 - 16:22:38
\begin{table}[!htbp] \centering 
  \caption{Baseline results revision} 
  \label{} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{The Dependent variable:} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Period}  & $-$0.055 & 0.010 &  &  &  &  \\ 
  & (0.051) & (0.081) &  &  & &  \\ 
  TCZ_c \times \text{Polluted}_i  &  & 0.107 &  &  &  &  \\ 
  &  & (0.094) & & &  &  \\ 
   \text{Period} \times \text{Polluted}_i  &  & $-$0.021 &  &  &  &  \\ 
  &  & (0.077) &  &  &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  &  & $-$0.123 &  &  &  &  \\ 
  &  & (0.090) &  &  &  &  \\ 
   TCZ_c \times \text{Polluted}_i \times \text{Period}  &  &  & $-$0.259 & $-$0.298$^{**}$ & $-$0.336$^{**}$ & $-$0.377$^{**}$ \\ 
  &  &  & (0.160) & (0.142) & (0.154) & (0.154) \\ 
   TCZ_c \times \text{count share SOE}_{i} \times \text{Period}  &  &  & $-$1.235$^{*}$ &  &  &  \\ 
  &  &  & (0.709) &  &  &  \\ 
   TCZ_c \times \text{Polluted}_i \times \text{count share SOE}_{i} \times \text{Period}  &  &  & 1.565$^{*}$ &  &  &  \\ 
  &  &  & (0.910) &  &  &  \\ 
   TCZ_c \times \text{output share SOE}_{i} \times \text{Period}  &  &  &  & $-$0.824 &  &  \\ 
  &  &  &  & (0.515) &  &  \\ 
   TCZ_c \times \text{Polluted}_i \times \text{output share SOE}_{i} \times \text{Period}  &  &  &  & 1.315$^{**}$ &  &  \\ 
  &  &  &  & (0.609) &  &  \\ 
   TCZ_c \times \text{capital share SOE}_{i} \times \text{Period}  &  &  &  &  & $-$0.682$^{*}$ &  \\ 
  &  &  &  &  & (0.399) &  \\ 
   TCZ_c \times \text{Polluted}_i \times \text{capital share SOE}_{i} \times \text{Period}  &  &  &  &  & 1.065$^{**}$ &  \\ 
  &  &  &  &  & (0.513) &  \\ 
   TCZ_c \times \text{labour share SOE}_{i} \times \text{Period}  &  &  &  &  &  & $-$1.020$^{**}$ \\ 
  &  &  &  &  &  & (0.518) \\ 
   TCZ_c \times \text{Polluted}_i \times \text{labour share SOE}_{i} \times \text{Period}  &  &  &  &  &  & 1.532$^{**}$ \\ 
  &  &  &  &  &  & (0.598) \\ 
 \hline \\[-1.8ex] 
 City fixed effects & Yes & Yes & No& No & No& No\\ 
Industry fixed effects & Yes & Yes & No& No & No& No\\ 
Year fixed effects & Yes & Yes & No& No & No& No\\ 
City-year fixed effects & No & No & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & Yes & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.345 & 0.345 & 0.851 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{6}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{6}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}