### 2020-04-27 Temp_tables/Temp_tables_revision/00_Archive/table_2_foreign.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[!htbp] \centering 
  \caption{Revision results: Foreing} 
  \label{} 
   \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{The Dependent variable:} \\ 
\cline{2-4} 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
  output_{cit} & $-$0.057 & $-$0.057 & $-$0.056 \\ 
  & (0.084) & (0.084) & (0.084) \\ 
  capital_{cit} & 0.929$^{***}$ & 0.929$^{***}$ & 0.935$^{***}$ \\ 
  & (0.356) & (0.355) & (0.357) \\ 
  labour_{cit} & 1.530$^{**}$ & 1.534$^{**}$ & 1.498$^{**}$ \\ 
  & (0.724) & (0.725) & (0.724) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.021 & $-$0.049 & 0.011 \\ 
  & (0.200) & (0.194) & (0.173) \\ 
   TCZ_c \times \text{Period} \times \text{output share Foreign}_{i} & 0.856 &  &  \\ 
  & (0.878) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share Foreign}_{i} & $-$0.874 &  &  \\ 
  & (1.558) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share Foreign}_{i}  &  & 0.650 &  \\ 
  &  & (0.794) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share Foreign}_{i}  &  & $-$0.555 &  \\ 
  &  & (1.338) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share Foreign}_{i}  &  &  & 1.437$^{*}$ \\ 
  &  &  & (0.830) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share Foreign}_{i}  &  &  & $-$1.061 \\ 
  &  &  & (1.583) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{3}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{3}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}