### 2020-04-27 Temp_tables/Temp_tables_revision/00_Archive/table_3_foreign_SOE.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

\begin{table}[!htbp] \centering 
  \caption{Revision results: Foreing -SOE} 
  \label{} 
  \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{The Dependent variable:} \\ 
\cline{2-4} 
\\[-1.8ex] & (1) & (2) & (3)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.294 & $-$0.435 & $-$0.342 \\ 
  & (0.238) & (0.287) & (0.234) \\ 
   TCZ_c \times \text{Period} \times \text{output share Foreign}_{i} & 0.464 &  &  \\ 
  & (0.963) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{output share SOE}_{i}  & $-$0.706 &  &  \\ 
  & (0.619) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share Foreign}_{i} & 0.073 &  &  \\ 
  & (1.607) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.278$^{*}$ &  &  \\ 
  & (0.748) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share Foreign}_{i}  &  & 0.082 &  \\ 
  &  & (0.910) &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share SOE}_{i}  &  & $-$0.661 &  \\ 
  &  & (0.503) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share Foreign}_{i}  &  & 0.627 &  \\ 
  &  & (1.544) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 1.176$^{*}$ &  \\ 
  &  & (0.670) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share Foreign}_{i}  &  &  & 0.937 \\ 
  &  &  & (0.937) \\ 
   TCZ_c \times \text{Period} \times \text{labour share SOE}_{i}  &  &  & $-$0.727 \\ 
  &  &  & (0.621) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share Foreign}_{i}  &  &  & 0.161 \\ 
  &  &  & (1.659) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 1.384$^{*}$ \\ 
  &  &  & (0.748) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{3}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{3}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{3}{l}{are clustered by city} \\ 
\end{tabular} 
\end{adjustbox}
\end{table}

\end{document}