### 2020-04-27 Temp_tables/Temp_tables_revision/01_Herfindhal/06_conc_ind_soe.tex

\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}

% Table created by stargazer v.5.2.2 by Marek Hlavac, Harvard University. E-mail: hlavac at fas.harvard.edu
% Date and time: Sun, Mar 08, 2020 - 12:20:15
\begin{table}[!htbp] \centering 
  \caption{Baseline results revision: concentration} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{The Dependent variable:} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.032 & $-$0.142 & $-$0.097 & 0.086 & 0.037 & 0.060 \\ 
  & (0.639) & (0.643) & (0.635) & (0.698) & (0.688) & (0.679) \\ 
   TCZ_c \times \text{Period} \times \text{Herfindahl}_{i} & 0.029 & 0.013 & 0.013 & 0.451 & 0.491 & 0.528 \\ 
  & (0.819) & (0.810) & (0.818) & (0.946) & (0.943) & (0.914) \\ 
   TCZ_c \times \text{Period} \times \text{output share SOE}_{i}  & $-$0.199 &  &  & $-$0.944$^{*}$ &  &  \\ 
  & (0.431) &  &  & (0.533) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share SOE}_{i}  &  & $-$0.103 &  &  & $-$0.774$^{*}$ &  \\ 
  &  & (0.325) &  &  & (0.408) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share SOE}_{i}  &  &  & $-$0.137 &  &  & $-$1.148$^{**}$ \\ 
  &  &  & (0.441) &  &  & (0.518) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{Herfindahl}_{i} & $-$0.404 & $-$0.240 & $-$0.345 & $-$0.597 & $-$0.582 & $-$0.687 \\ 
  & (0.932) & (0.937) & (0.931) & (1.034) & (1.019) & (1.001) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.216$^{**}$ &  &  & 1.478$^{**}$ &  &  \\ 
  & (0.586) &  &  & (0.649) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 0.787$^{*}$ &  &  & 1.175$^{**}$ &  \\ 
  &  & (0.459) &  &  & (0.541) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 1.048$^{*}$ &  &  & 1.702$^{***}$ \\ 
  &  &  & (0.572) &  &  & (0.612) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.346 & 0.346 & 0.346 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{6}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{6}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{6}{l}{are clustered by city} \\ 
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}