### 2020-04-27 Temp_tables/Temp_tables_revision/02_table_7_improvement/Decile/020_capital_industry_thresh.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Sun, Mar 29, 2020 - 11:17:52
\begin{table}[!htbp] \centering 
  \caption{Deciles capital} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{9}{c}{Dependent variable: \text { SO2 emission }_{i k t}} \\ 
\cline{2-10} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.409 & $-$0.894$^{**}$ & 1.151 & $-$0.768$^{*}$ & $-$0.134 & $-$0.560 & 0.121 & $-$0.076 & 0.910$^{***}$ \\ 
  & (0.569) & (0.349) & (0.820) & (0.437) & (0.388) & (0.436) & (0.274) & (0.611) & (0.337) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 1,503 & 2,377 & 2,195 & 3,028 & 4,428 & 4,126 & 3,630 & 2,643 & 3,768 \\ 
R$^{2}$ & 0.925 & 0.901 & 0.903 & 0.884 & 0.896 & 0.885 & 0.920 & 0.908 & 0.888 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank share of [output, capital, employement] of the SOEs \
More specifically, the low deciles means a low presence of SOEs firms in the industries. \
Deciles close to one, however, implies a stronger share of [output, capital, employement] in the \
industries belong to those deciles.
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\%
}

\end{tablenotes}
\end{table}

\end{document}