### 2020-04-27 Temp_tables/Temp_tables_revision/02_table_7_improvement/Decile/021_capital_industry_cum.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Sun, Mar 29, 2020 - 11:19:11
\begin{table}[!htbp] \centering 
  \caption{Cumulative Deciles output} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{9}{c}{Dependent variable: \text { SO2 emission }_{i k t}} \\ 
\cline{2-10} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.409 & $-$0.803$^{***}$ & $-$0.563 & $-$0.544$^{**}$ & $-$0.332$^{*}$ & $-$0.317$^{**}$ & $-$0.272$^{**}$ & $-$0.248$^{*}$ & $-$0.136 \\ 
  & (0.569) & (0.263) & (0.350) & (0.245) & (0.173) & (0.150) & (0.131) & (0.130) & (0.127) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 1,503 & 3,880 & 6,075 & 9,103 & 13,531 & 17,657 & 21,287 & 23,930 & 27,698 \\ 
R$^{2}$ & 0.925 & 0.889 & 0.871 & 0.855 & 0.858 & 0.852 & 0.856 & 0.855 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank share of [output, capital, employement] of the SOEs \
More specifically, the low deciles means a low presence of SOEs firms in the industries. \
Deciles close to one, however, implies a stronger share of [output, capital, employement] in the \
industries belong to those deciles.
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\%
}

\end{tablenotes}
\end{table}

\end{document}