### 2020-04-27 Temp_tables/Temp_tables_revision/02_table_7_improvement/wrong_cut_bin/010_SOE_city_industry.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Fri, Mar 27, 2020 - 09:33:03
\begin{table}[!htbp] \centering 
  \caption{SOE dominating city-sectors} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{12}{c}{The Dependent variable:} \\ 
\cline{2-13} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12)\\
 \\[-1.8ex]& decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.038 & $-$0.267 & 0.101 & 0.373 & $-$0.166 & $-$0.197 & 0.031 & $-$0.105 & $-$0.027 & $-$0.299 & 0.012 & $-$1.184 \\ 
  & (0.255) & (0.440) & (0.450) & (1.162) & (0.217) & (0.368) & (0.348) & (1.034) & (0.267) & (0.441) & (0.442) & (1.025) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 8,840 & 4,946 & 4,248 & 1,842 & 8,835 & 5,032 & 5,402 & 1,821 & 8,664 & 4,782 & 4,790 & 1,780 \\ 
R$^{2}$ & 0.879 & 0.906 & 0.923 & 0.966 & 0.883 & 0.909 & 0.912 & 0.964 & 0.883 & 0.913 & 0.918 & 0.960 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{12}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{12}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{12}{l}{are clustered by city} \\ 
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}