### 2020-04-27 Temp_tables/Temp_tables_revision/02_table_7_improvement/wrong_cut_bin/020_SOE_industry.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Fri, Mar 27, 2020 - 09:39:00
\begin{table}[!htbp] \centering 
  \caption{SOE dominating sectors} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{12}{c}{The Dependent variable:} \\ 
\cline{2-13} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12)\\
 \\[-1.8ex]& decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.059 & 0.031 & 0.191 & 0.191 & $-$0.080 & 0.074 & 0.161 & 0.289 & $-$0.081 & $-$0.009 & 0.273 & 0.151 \\ 
  & (0.134) & (0.160) & (0.183) & (0.183) & (0.140) & (0.172) & (0.184) & (0.218) & (0.142) & (0.160) & (0.203) & (0.237) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 24,942 & 18,946 & 10,055 & 10,027 & 24,601 & 17,259 & 11,813 & 9,459 & 24,388 & 18,790 & 11,535 & 8,218 \\ 
R$^{2}$ & 0.852 & 0.859 & 0.869 & 0.868 & 0.852 & 0.858 & 0.868 & 0.868 & 0.851 & 0.857 & 0.861 & 0.872 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{12}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{12}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{12}{l}{are clustered by city} \\ 
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}