### 2020-04-27 Temp_tables/Temp_tables_revision/02_table_7_improvement/wrong_cut_bin/021_PRIVATE_industry.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Fri, Mar 27, 2020 - 09:39:54
\begin{table}[!htbp] \centering 
  \caption{Private dominating sectors} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{12}{c}{The Dependent variable:} \\ 
\cline{2-13} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12)\\
 \\[-1.8ex]& decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.787$^{**}$ & $-$0.432$^{**}$ & $-$0.233$^{*}$ & $-$0.233$^{*}$ & $-$0.563 & $-$0.343$^{**}$ & $-$0.306$^{**}$ & $-$0.278$^{**}$ & $-$0.541$^{*}$ & $-$0.254 & $-$0.305$^{**}$ & $-$0.204 \\ 
  & (0.326) & (0.215) & (0.136) & (0.136) & (0.350) & (0.174) & (0.137) & (0.131) & (0.299) & (0.251) & (0.137) & (0.140) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 5,734 & 11,730 & 20,621 & 20,649 & 6,075 & 13,417 & 18,863 & 21,217 & 6,288 & 11,886 & 19,141 & 22,458 \\ 
R$^{2}$ & 0.876 & 0.849 & 0.855 & 0.855 & 0.871 & 0.859 & 0.851 & 0.856 & 0.875 & 0.847 & 0.860 & 0.854 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{12}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{12}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{12}{l}{are clustered by city} \\ 
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}