### 2020-04-27 Temp_tables/Temp_tables_revision/02_table_7_improvement/wrong_cut_bin/030_SOE_city.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Fri, Mar 27, 2020 - 09:43:16
\begin{table}[!htbp] \centering 
  \caption{SOE dominating city} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{12}{c}{The Dependent variable:} \\ 
\cline{2-13} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12)\\
 \\[-1.8ex]& decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7 & decile .3 & median &  mean & decile .7\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.187 & $-$0.283 & $-$0.247 & $-$0.068 & $-$0.084 & $-$0.225 & $-$0.205 & $-$0.032 & $-$0.146 & $-$0.266 & $-$0.236 & $-$0.206 \\ 
  & (0.160) & (0.174) & (0.208) & (0.247) & (0.155) & (0.172) & (0.174) & (0.270) & (0.155) & (0.175) & (0.184) & (0.258) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 19,884 & 13,240 & 10,516 & 7,465 & 20,004 & 12,942 & 13,062 & 7,528 & 19,730 & 13,214 & 12,226 & 6,668 \\ 
R$^{2}$ & 0.858 & 0.866 & 0.867 & 0.877 & 0.858 & 0.867 & 0.867 & 0.875 & 0.857 & 0.862 & 0.864 & 0.884 \\ 
\hline 
\hline \\[-1.8ex] 
\textit{Note:}  & \multicolumn{12}{l}{$^{*}$p$<$0.1; $^{**}$p$<$0.05; $^{***}$p$<$0.01} \\ 
 & \multicolumn{12}{l}{Heteroskedasticity-robust standard errors in parentheses} \\ 
 & \multicolumn{12}{l}{are clustered by city} \\ 
\end{tabular}
\end{adjustbox}
\end{table}

\end{document}