### 2020-04-27 Temp_tables/Temp_tables_revision/03_Paper_version_Revised/02_table_5_rob_1.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sat, Apr 04, 2020 - 15:06:11
\begin{table}[!htbp] \centering 
  \caption{Soft budget constraint as reflected in the SOEs reaction to the  SPZc and coastal area's policies in China } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.314$^{**}$ & $-$0.307$^{**}$ & $-$0.320$^{**}$ & $-$0.320$^{**}$ & $-$0.331$^{**}$ & $-$0.378$^{**}$ \\ 
  & (0.129) & (0.140) & (0.141) & (0.141) & (0.155) & (0.156) \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i  & 0.114 & 0.077 & 0.067 & 0.070 & 0.004 & 0.019 \\ 
  & (0.098) & (0.104) & (0.109) & (0.112) & (0.120) & (0.124) \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i  & $-$0.182 & $-$0.159 & $-$0.180 & $-$0.034 & $-$0.034 & $-$0.044 \\ 
  & (0.113) & (0.122) & (0.125) & (0.133) & (0.139) & (0.147) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.158$^{**}$ &  &  & 1.305$^{*}$ &  &  \\ 
  & (0.533) &  &  & (0.672) &  &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & $-$0.122 &  &  & 0.300 &  &  \\ 
  & (0.518) &  &  & (0.608) &  &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & $-$0.100 &  &  & $-$0.327 &  &  \\ 
  & (0.513) &  &  & (0.610) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 0.695 &  &  & 0.924$^{*}$ &  \\ 
  &  & (0.438) &  &  & (0.534) &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 0.127 &  &  & 0.477 &  \\ 
  &  & (0.398) &  &  & (0.438) &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & $-$0.105 &  &  & $-$0.150 &  \\ 
  &  & (0.402) &  &  & (0.473) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 0.917 &  &  & 1.391$^{**}$ \\ 
  &  &  & (0.564) &  &  & (0.656) \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 0.151 &  &  & 0.525 \\ 
  &  &  & (0.506) &  &  & (0.558) \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & $-$0.108 &  &  & $-$0.142 \\ 
  &  &  & (0.488) &  &  & (0.588) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.348 & 0.347 & 0.348 & 0.851 & 0.851 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}