### 2020-04-27 Temp_tables/Temp_tables_revision/03_Paper_version_Revised/03_table_6_rob_2.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Sat, Apr 04, 2020 - 10:22:58
\begin{table}[!htbp] \centering 
  \caption{Environmental policy,Wealth and Population pressure } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable: \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   \text{Polluted}_i \times \text{ln gdp per cap)}_{ct}  & $-$0.042 & $-$0.042 & $-$0.042 & $-$0.034 & $-$0.029 & $-$0.036 \\ 
  & (0.065) & (0.065) & (0.065) & (0.269) & (0.269) & (0.269) \\ 
   \text{Polluted}_i \times \text{ln population)}_{ct}  & 0.041 & 0.041 & 0.041 & 0.170 & 0.172 & 0.172 \\ 
  & (0.050) & (0.050) & (0.050) & (0.253) & (0.253) & (0.253) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.294$^{**}$ & $-$0.296$^{**}$ & $-$0.315$^{**}$ & $-$0.280$^{*}$ & $-$0.317$^{**}$ & $-$0.360$^{**}$ \\ 
  & (0.127) & (0.137) & (0.138) & (0.144) & (0.156) & (0.156) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.152$^{**}$ &  &  & 1.327$^{**}$ &  &  \\ 
  & (0.539) &  &  & (0.605) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 0.773$^{*}$ &  &  & 1.074$^{**}$ &  \\ 
  &  & (0.432) &  &  & (0.510) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 0.998$^{*}$ &  &  & 1.545$^{***}$ \\ 
  &  &  & (0.529) &  &  & (0.594) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,195 & 30,195 & 30,195 & 30,195 & 30,195 & 30,195 \\ 
R$^{2}$ & 0.345 & 0.345 & 0.345 & 0.852 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}