### 2020-04-27 Temp_tables/Temp_tables_revision/03_Paper_version_Revised/04_table_7_rob_3.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Sat, Apr 04, 2020 - 11:27:39
\begin{table}[!htbp] \centering 
  \caption{Presence of Foreign firms} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable: \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.101 & $-$0.101 & $-$0.051 & 0.078 & 0.123 & 0.044 \\ 
  & (0.169) & (0.164) & (0.156) & (0.211) & (0.203) & (0.180) \\ 
   TCZ_c \times \text{Period} \times \text{output share Foreign}_{i} & $-$0.387 &  &  & 0.937 &  &  \\ 
  & (0.662) &  &  & (0.919) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share Foreign}_{i} & $-$0.299 &  &  & $-$1.448 &  &  \\ 
  & (1.019) &  &  & (1.237) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share Foreign}_{i}  &  & $-$0.291 &  &  & 0.984 &  \\ 
  &  & (0.620) &  &  & (0.781) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share Foreign}_{i}  &  & $-$0.249 &  &  & $-$1.710 &  \\ 
  &  & (0.942) &  &  & (1.109) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share Foreign}_{i}  &  &  & 0.004 &  &  & 1.203 \\ 
  &  &  & (0.599) &  &  & (0.757) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share Foreign}_{i}  &  &  & $-$0.798 &  &  & $-$1.172 \\ 
  &  &  & (1.162) &  &  & (1.167) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.346 & 0.346 & 0.346 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}