### 2020-04-27 Temp_tables/Temp_tables_revision/03_Paper_version_Revised/07_table_9_size_capital.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time: Sat, Apr 04, 2020 - 13:38:12
\begin{table}[!htbp] \centering 
  \caption{Deciles capital} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable: \text { SO2 emission }_{i k t}} \\ 
\cline{2-11} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9 &  Baseline\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.610 & $-$0.342 & $-$0.467$^{**}$ & $-$0.467$^{**}$ & $-$0.406$^{**}$ & $-$0.401$^{**}$ & $-$0.368$^{**}$ & $-$0.337$^{**}$ & $-$0.343$^{**}$ & $-$0.336$^{**}$ \\ 
  & (0.533) & (0.253) & (0.213) & (0.202) & (0.190) & (0.168) & (0.155) & (0.154) & (0.154) & (0.154) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  & 2.183 & 1.785 & 1.380 & 1.545$^{*}$ & 1.289$^{*}$ & 1.245$^{*}$ & 1.192$^{**}$ & 1.027$^{*}$ & 1.073$^{**}$ & 1.065$^{**}$ \\ 
  & (2.064) & (1.099) & (0.844) & (0.827) & (0.732) & (0.656) & (0.548) & (0.529) & (0.515) & (0.513) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 7,646 & 13,563 & 17,682 & 22,109 & 25,418 & 28,062 & 29,422 & 30,132 & 30,388 & 30,676 \\ 
R$^{2}$ & 0.877 & 0.860 & 0.853 & 0.845 & 0.846 & 0.847 & 0.848 & 0.850 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank of the city-industry concentration (Herfhindal index) \
More specifically, the low deciles means a low concentration by city-industry \
Deciles close to one, however, implies a stronger concentration.
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}