### 2020-04-27 Temp_tables/Temp_tables_revision/03_Paper_version_Revised/10_decile_7.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sat, Apr 04, 2020 - 14:51:25
\begin{table}[!htbp] \centering 
  \caption{Decile 7} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
  output_{cit} & 0.136 & 0.136 & 0.137 & $-$0.059 & $-$0.059 & $-$0.059 \\ 
  & (0.154) & (0.154) & (0.154) & (0.114) & (0.114) & (0.113) \\ 
  capital_{cit} & 1.511$^{**}$ & 1.513$^{**}$ & 1.511$^{**}$ & 0.929$^{*}$ & 0.933$^{**}$ & 0.937$^{**}$ \\ 
  & (0.733) & (0.735) & (0.735) & (0.476) & (0.475) & (0.474) \\ 
  labour_{cit} & 3.038$^{***}$ & 3.041$^{***}$ & 3.035$^{***}$ & 1.550 & 1.545 & 1.538 \\ 
  & (0.829) & (0.831) & (0.829) & (1.023) & (1.023) & (1.021) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} & $-$0.304$^{**}$ & $-$0.305$^{**}$ & $-$0.324$^{**}$ & $-$0.298$^{**}$ & $-$0.336$^{**}$ & $-$0.377$^{**}$ \\ 
  & (0.124) & (0.135) & (0.136) & (0.142) & (0.154) & (0.154) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} \times \text{output share SOE}_{i}  & 1.127$^{**}$ &  &  & 1.315$^{**}$ &  &  \\ 
  & (0.533) &  &  & (0.609) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} \times \text{capital share SOE}_{i}  &  & 0.744$^{*}$ &  &  & 1.065$^{**}$ &  \\ 
  &  & (0.431) &  &  & (0.513) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} \times \text{labour share SOE}_{i}  &  &  & 0.958$^{*}$ &  &  & 1.532$^{**}$ \\ 
  &  &  & (0.528) &  &  & (0.598) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Year fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.346 & 0.346 & 0.346 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}