### 2020-04-27 Temp_tables/Temp_tables_revision/03_Paper_version_Revised/12_decile_9.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sat, Apr 04, 2020 - 14:57:03
\begin{table}[!htbp] \centering 
  \caption{Decile 9} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
  output_{cit} & 0.142 & 0.140 & 0.142 & $-$0.058 & $-$0.058 & $-$0.058 \\ 
  & (0.156) & (0.155) & (0.156) & (0.115) & (0.115) & (0.115) \\ 
  capital_{cit} & 1.501$^{**}$ & 1.501$^{**}$ & 1.503$^{**}$ & 0.923$^{*}$ & 0.924$^{*}$ & 0.924$^{*}$ \\ 
  & (0.735) & (0.735) & (0.735) & (0.477) & (0.476) & (0.476) \\ 
  labour_{cit} & 3.035$^{***}$ & 3.046$^{***}$ & 3.032$^{***}$ & 1.573 & 1.571 & 1.571 \\ 
  & (0.828) & (0.830) & (0.828) & (1.016) & (1.015) & (1.014) \\ 
   TCZ_c \times \text{Period} :decile\_so2\_9Above & $-$0.496$^{***}$ & $-$0.470$^{***}$ & $-$0.473$^{***}$ & $-$0.148 & $-$0.124 & $-$0.096 \\ 
  & (0.156) & (0.161) & (0.178) & (0.172) & (0.190) & (0.202) \\ 
   TCZ_c \times \text{Period} :decile\_so2\_9Above \times \text{output share SOE}_{i}  & 1.587$^{**}$ &  &  & 0.548 &  &  \\ 
  & (0.631) &  &  & (0.840) &  &  \\ 
   TCZ_c \times \text{Period} :decile\_so2\_9Above \times \text{capital share SOE}_{i}  &  & 0.883$^{*}$ &  &  & 0.286 &  \\ 
  &  & (0.528) &  &  & (0.654) &  \\ 
   TCZ_c \times \text{Period} :decile\_so2\_9Above \times \text{labour share SOE}_{i}  &  &  & 1.224$^{**}$ &  &  & 0.227 \\ 
  &  &  & (0.620) &  &  & (0.739) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Year fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.345 & 0.345 & 0.345 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}