### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/02_table_5_rob_1.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 09:47:02
\begin{table}[!htbp] \centering 
  \caption{Soft budget constraint as reflected in the SOEs reaction to the  SPZ s and coastal area's policies in China } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.109 & $-$0.104 & $-$0.107 & $-$0.227$^{*}$ & $-$0.246$^{*}$ & $-$0.228$^{*}$ \\ 
  & (0.112) & (0.119) & (0.117) & (0.128) & (0.133) & (0.129) \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i  & 0.012 & $-$0.021 & 0.002 & 0.068 & 0.060 & 0.066 \\ 
  & (0.081) & (0.088) & (0.087) & (0.098) & (0.107) & (0.108) \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i  & $-$0.147 & $-$0.151 & $-$0.155 & $-$0.056 & $-$0.056 & $-$0.061 \\ 
  & (0.095) & (0.098) & (0.097) & (0.109) & (0.113) & (0.113) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci}  & $-$0.221 &  &  & 0.497 &  &  \\ 
  & (0.319) &  &  & (0.377) &  &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci}  & 0.460 &  &  & 0.166 &  &  \\ 
  & (0.281) &  &  & (0.395) &  &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci}  & 0.267 &  &  & 0.117 &  &  \\ 
  & (0.283) &  &  & (0.338) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{ci}  &  & $-$0.174 &  &  & 0.439 &  \\ 
  &  & (0.283) &  &  & (0.332) &  \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{ci}  &  & 0.524$^{**}$ &  &  & 0.143 &  \\ 
  &  & (0.247) &  &  & (0.358) &  \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{ci}  &  & 0.206 &  &  & 0.079 &  \\ 
  &  & (0.243) &  &  & (0.304) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{ci}  &  &  & $-$0.193 &  &  & 0.384 \\ 
  &  &  & (0.305) &  &  & (0.360) \\ 
   SPZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{ci}  &  &  & 0.423 &  &  & 0.109 \\ 
  &  &  & (0.267) &  &  & (0.379) \\ 
   Coastal_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{ci}  &  &  & 0.276 &  &  & 0.127 \\ 
  &  &  & (0.259) &  &  & (0.328) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.348 & 0.348 & 0.348 & 0.852 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}