### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/03_table_6_rob_2.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 09:50:45
\begin{table}[!htbp] \centering 
  \caption{Environmental policy,Wealth and Population pressure } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   \text{Polluted}_i \times \text{ln gdp per cap)}_{ct}  & $-$0.025 & $-$0.027 & $-$0.024 & $-$0.021 & $-$0.028 & $-$0.023 \\ 
  & (0.066) & (0.066) & (0.066) & (0.268) & (0.268) & (0.268) \\ 
   \text{Polluted}_i \times \text{ln population)}_{ct}  & 0.033 & 0.031 & 0.028 & 0.173 & 0.173 & 0.175 \\ 
  & (0.050) & (0.051) & (0.051) & (0.252) & (0.252) & (0.251) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.114 & $-$0.126 & $-$0.120 & $-$0.196 & $-$0.220 & $-$0.200 \\ 
  & (0.111) & (0.117) & (0.115) & (0.137) & (0.139) & (0.138) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci}  & $-$0.058 &  &  & 0.537 &  &  \\ 
  & (0.328) &  &  & (0.376) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{ci}  &  & 0.031 &  &  & 0.503 &  \\ 
  &  & (0.283) &  &  & (0.328) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{ci}  &  &  & $-$0.013 &  &  & 0.431 \\ 
  &  &  & (0.308) &  &  & (0.355) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,195 & 30,195 & 30,195 & 30,195 & 30,195 & 30,195 \\ 
R$^{2}$ & 0.346 & 0.346 & 0.346 & 0.852 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}