### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/04bis_table_7_rob_3.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 10:34:36
\begin{table}[!htbp] \centering 
  \caption{Presence of Foreign firms} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.422$^{*}$ & $-$0.521$^{*}$ & $-$0.356 & $-$0.186 & $-$0.274 & $-$0.347 \\ 
  & (0.228) & (0.288) & (0.258) & (0.296) & (0.347) & (0.305) \\ 
   TCZ_c \times \text{Period} \times \text{output share Foreign}_{i} & $-$0.647 &  &  & 0.326 &  &  \\ 
  & (0.756) &  &  & (1.090) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{output share SOE}_{ci} \_ & $-$0.377 &  &  & $-$0.844 &  &  \\ 
  & (0.466) &  &  & (0.597) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share Foreign}_{i} & 0.719 &  &  & $-$0.626 &  &  \\ 
  & (1.097) &  &  & (1.406) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci} \_ & 1.069$^{*}$ &  &  & 1.030 &  &  \\ 
  & (0.568) &  &  & (0.673) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share Foreign}_{i}  &  & $-$0.525 &  &  & 0.109 &  \\ 
  &  & (0.805) &  &  & (0.995) &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share SOE}_{ci} \_ &  & $-$0.209 &  &  & $-$0.751 &  \\ 
  &  & (0.406) &  &  & (0.495) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share Foreign}_{i}  &  & 1.094 &  &  & $-$0.527 &  \\ 
  &  & (1.156) &  &  & (1.401) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{ci} \_ &  & 0.827 &  &  & 0.918 &  \\ 
  &  & (0.536) &  &  & (0.606) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share Foreign}_{i}  &  &  & $-$0.129 &  &  & 0.475 \\ 
  &  &  & (0.737) &  &  & (0.983) \\ 
   TCZ_c \times \text{Period} \times \text{labour share SOE}_{ci} \_ &  &  & $-$0.177 &  &  & $-$0.843 \\ 
  &  &  & (0.455) &  &  & (0.601) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share Foreign}_{i}  &  &  & 0.230 &  &  & 0.014 \\ 
  &  &  & (1.314) &  &  & (1.399) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{ci} \_ &  &  & 0.778 &  &  & 1.154$^{*}$ \\ 
  &  &  & (0.567) &  &  & (0.664) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\
R$^{2}$ & 0.346 & 0.346 & 0.346 & 0.851 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}