### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/05_table_8_rob_4.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 10:37:00
\begin{table}[!htbp] \centering 
  \caption{Concentration of industry} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.215 & $-$0.219 & $-$0.198 & $-$0.125 & $-$0.156 & $-$0.125 \\ 
  & (0.648) & (0.647) & (0.647) & (0.657) & (0.661) & (0.660) \\ 
   TCZ_c \times \text{Period} \times \text{Herfindahl}_{i} & $-$0.060 & $-$0.036 & $-$0.029 & 0.248 & 0.232 & 0.246 \\ 
  & (0.779) & (0.768) & (0.773) & (0.867) & (0.875) & (0.875) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{Herfindahl}_{i} & 0.137 & 0.128 & 0.108 & $-$0.136 & $-$0.123 & $-$0.138 \\ 
  & (0.902) & (0.898) & (0.900) & (0.924) & (0.931) & (0.930) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci}  & $-$0.077 &  &  & 0.533 &  &  \\ 
  & (0.323) &  &  & (0.373) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{ci}  &  & 0.007 &  &  & 0.485 &  \\ 
  &  & (0.277) &  &  & (0.324) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{ci}  &  &  & $-$0.048 &  &  & 0.416 \\ 
  &  &  & (0.303) &  &  & (0.354) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Yes fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.347 & 0.347 & 0.347 & 0.851 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}