### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/06_table_9_size_output.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 10:38:08
\begin{table}[!htbp] \centering 
  \caption{Deciles output} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-11} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9 &  Baseline\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.348 & $-$0.127 & $-$0.280 & $-$0.201 & $-$0.228 & $-$0.252$^{*}$ & $-$0.227$^{*}$ & $-$0.210 & $-$0.221$^{*}$ & $-$0.213 \\ 
  & (0.357) & (0.183) & (0.177) & (0.163) & (0.156) & (0.145) & (0.136) & (0.135) & (0.134) & (0.135) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{ci}  & 2.118$^{*}$ & 0.628 & 0.535 & 0.148 & 0.482 & 0.608 & 0.546 & 0.478 & 0.541 & 0.529 \\ 
  & (1.118) & (0.798) & (0.500) & (0.572) & (0.488) & (0.427) & (0.394) & (0.383) & (0.373) & (0.374) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 7,646 & 13,563 & 17,682 & 22,109 & 25,418 & 28,062 & 29,422 & 30,132 & 30,388 & 30,676 \\ 
R$^{2}$ & 0.878 & 0.860 & 0.853 & 0.845 & 0.846 & 0.847 & 0.848 & 0.850 & 0.851 & 0.851 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank of the city-industry concentration (Herfhindal index) \
More specifically, the low deciles means a low concentration by city-industry \
Deciles close to one, however, implies a stronger concentration.
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}