### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/10_decile_7.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 09:43:17
\begin{table}[!htbp] \centering 
  \caption{Decile 7} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
  output_{cit} & 0.147 & 0.147 & 0.147 & $-$0.060 & $-$0.061 & $-$0.061 \\ 
  & (0.160) & (0.159) & (0.159) & (0.114) & (0.113) & (0.112) \\ 
  capital_{cit} & 1.468$^{**}$ & 1.468$^{**}$ & 1.468$^{**}$ & 0.945$^{**}$ & 0.946$^{**}$ & 0.949$^{**}$ \\ 
  & (0.723) & (0.727) & (0.722) & (0.474) & (0.474) & (0.471) \\ 
  labour_{cit} & 2.993$^{***}$ & 2.991$^{***}$ & 2.990$^{***}$ & 1.513 & 1.499 & 1.495 \\ 
  & (0.821) & (0.819) & (0.822) & (1.006) & (1.010) & (1.002) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} & $-$0.121 & $-$0.131 & $-$0.124 & $-$0.213 & $-$0.235$^{*}$ & $-$0.215 \\ 
  & (0.110) & (0.116) & (0.114) & (0.135) & (0.137) & (0.136) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} \times \text{output share SOE}_{ci}  & $-$0.093 &  &  & 0.529 &  &  \\ 
  & (0.327) &  &  & (0.374) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} \times \text{capital share SOE}_{ci}  &  & $-$0.008 &  &  & 0.485 &  \\ 
  &  & (0.282) &  &  & (0.324) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 7}_{i} \times \text{labour share SOE}_{ci}  &  &  & $-$0.059 &  &  & 0.415 \\ 
  &  &  & (0.307) &  &  & (0.355) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Year fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.347 & 0.347 & 0.347 & 0.851 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}