### 2020-04-27 Temp_tables/Temp_tables_revision/04_city_industry/11_decile_8.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Sun, Apr 05, 2020 - 09:43:29
\begin{table}[!htbp] \centering 
  \caption{Decile 8} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
  output_{cit} & 0.149 & 0.149 & 0.148 & $-$0.059 & $-$0.060 & $-$0.061 \\ 
  & (0.159) & (0.159) & (0.159) & (0.114) & (0.113) & (0.112) \\ 
  capital_{cit} & 1.437$^{**}$ & 1.442$^{**}$ & 1.442$^{**}$ & 0.936$^{**}$ & 0.937$^{**}$ & 0.941$^{**}$ \\ 
  & (0.729) & (0.729) & (0.727) & (0.468) & (0.468) & (0.465) \\ 
  labour_{cit} & 3.017$^{***}$ & 3.011$^{***}$ & 3.016$^{***}$ & 1.510 & 1.497 & 1.494 \\ 
  & (0.820) & (0.818) & (0.821) & (1.004) & (1.006) & (0.999) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 8}_{i} & $-$0.183$^{*}$ & $-$0.201$^{*}$ & $-$0.199$^{*}$ & $-$0.226$^{*}$ & $-$0.236$^{*}$ & $-$0.229$^{*}$ \\ 
  & (0.107) & (0.113) & (0.110) & (0.133) & (0.135) & (0.134) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 8}_{i} \times \text{output share SOE}_{ci}  & 0.065 &  &  & 0.476 &  &  \\ 
  & (0.332) &  &  & (0.376) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 8}_{i} \times \text{capital share SOE}_{ci}  &  & 0.130 &  &  & 0.408 &  \\ 
  &  & (0.294) &  &  & (0.317) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted decile 8}_{i} \times \text{labour share SOE}_{ci}  &  &  & 0.137 &  &  & 0.397 \\ 
  &  &  & (0.316) &  &  & (0.353) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & No & No & No \\ 
Industry fixed effects & Yes & Yes & Yes & No & No & No \\ 
Year fixed effects & Yes & Yes & Yes & No & No & No \\ 
City-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
Industry-year fixed effects & No & No & No & Yes & Yes & Yes \\ 
City-industry fixed effects & No & No & No & Yes & Yes & Yes \\ 
Observations & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 & 30,676 \\ 
R$^{2}$ & 0.346 & 0.347 & 0.346 & 0.851 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}