### 2020-04-27 Temp_tables/Temp_tables_revision/05_Kuznet/01_kuznet.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Mon, Apr 06, 2020 - 19:48:59
\begin{table}[!htbp] \centering 
  \caption{Test for the Kuznet curve } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   \text{(ln gdp per cap)}_{ct}  & 2.365$^{***}$ & 2.363$^{***}$ & 2.383$^{***}$ & 3.155$^{***}$ & 3.162$^{***}$ & 3.173$^{***}$ \\ 
  & (0.751) & (0.752) & (0.750) & (0.931) & (0.933) & (0.929) \\ 
   \text{(ln gdp per cap) squared}_{ct}  & $-$0.112$^{***}$ & $-$0.112$^{***}$ & $-$0.113$^{***}$ & $-$0.151$^{***}$ & $-$0.151$^{***}$ & $-$0.152$^{***}$ \\ 
  & (0.037) & (0.037) & (0.037) & (0.045) & (0.045) & (0.045) \\ 
   \text{(ln population)}_{ct}  & 0.042 & 0.042 & 0.042 & 0.043 & 0.043 & 0.043 \\ 
  & (0.089) & (0.088) & (0.088) & (0.088) & (0.088) & (0.088) \\ 
   \text{(ln gdp per cap)}_{ct} \times \text{Output SOE median)}  &  &  &  & $-$2.734$^{*}$ & $-$2.760$^{*}$ & $-$2.739$^{*}$ \\ 
  &  &  &  & (1.604) & (1.603) & (1.602) \\ 
   \text{(ln gdp per cap) squared}_{ct} \times \text{Output SOE median)}  &  &  &  & 0.135$^{*}$ & 0.137$^{*}$ & 0.136$^{*}$ \\ 
  &  &  &  & (0.079) & (0.079) & (0.079) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.303$^{**}$ & $-$0.306$^{**}$ & $-$0.325$^{**}$ & $-$0.307$^{**}$ & $-$0.310$^{**}$ & $-$0.330$^{**}$ \\ 
  & (0.127) & (0.137) & (0.138) & (0.128) & (0.138) & (0.139) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.160$^{**}$ &  &  & 1.170$^{**}$ &  &  \\ 
  & (0.536) &  &  & (0.537) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 0.779$^{*}$ &  &  & 0.785$^{*}$ &  \\ 
  &  & (0.430) &  &  & (0.429) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 1.011$^{*}$ &  &  & 1.021$^{*}$ \\ 
  &  &  & (0.527) &  &  & (0.527) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 30,195 & 30,195 & 30,195 & 30,195 & 30,195 & 30,195 \\ 
R$^{2}$ & 0.345 & 0.345 & 0.345 & 0.345 & 0.345 & 0.345 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-6).
Output SOE median is a dummy variable taking the value of 1 if the output share of the SOE computed \
at the city level is above the median.
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}