### 2020-04-27 Temp_tables/Temp_tables_revision/05_Kuznet/02_kuznet_decile_output.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Tue, Apr 07, 2020 - 12:06:50
\begin{table}[!htbp] \centering 
  \caption{Deciles output} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-11} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9 &  Baseline\\
 \hline \\[-1.8ex] 
   \text{(ln gdp per cap)}_{ct}  & 4.469$^{**}$ & 2.962$^{*}$ & 2.494$^{**}$ & 2.822$^{**}$ & 3.092$^{***}$ & 3.078$^{***}$ & 2.869$^{***}$ & 2.457$^{***}$ & 2.571$^{***}$ & 2.365$^{***}$ \\ 
  & (2.070) & (1.685) & (1.211) & (1.121) & (0.990) & (0.885) & (0.850) & (0.790) & (0.765) & (0.751) \\ 
   \text{(ln gdp per cap) squared}_{ct}  & $-$0.219$^{**}$ & $-$0.143$^{*}$ & $-$0.120$^{**}$ & $-$0.140$^{**}$ & $-$0.156$^{***}$ & $-$0.151$^{***}$ & $-$0.135$^{***}$ & $-$0.118$^{***}$ & $-$0.121$^{***}$ & $-$0.112$^{***}$ \\ 
  & (0.098) & (0.081) & (0.060) & (0.055) & (0.048) & (0.043) & (0.041) & (0.038) & (0.037) & (0.037) \\ 
   \text{(ln population)}_{ct}  & 0.019 & 0.009 & 0.081 & 0.060 & $-$0.00005 & 0.041 & 0.059 & 0.051 & 0.047 & 0.042 \\ 
  & (0.140) & (0.117) & (0.113) & (0.110) & (0.103) & (0.098) & (0.096) & (0.092) & (0.092) & (0.089) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$1.812$^{***}$ & $-$0.531$^{*}$ & $-$0.185 & $-$0.359$^{*}$ & $-$0.366$^{**}$ & $-$0.391$^{**}$ & $-$0.336$^{**}$ & $-$0.356$^{**}$ & $-$0.321$^{**}$ & $-$0.303$^{**}$ \\ 
  & (0.623) & (0.310) & (0.231) & (0.201) & (0.186) & (0.166) & (0.153) & (0.146) & (0.135) & (0.127) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 7.612 & 2.475 & 1.662 & 1.829$^{*}$ & 2.182$^{**}$ & 2.197$^{***}$ & 1.760$^{**}$ & 1.747$^{**}$ & 1.312$^{**}$ & 1.160$^{**}$ \\ 
  & (4.919) & (2.225) & (1.203) & (1.073) & (0.963) & (0.838) & (0.791) & (0.702) & (0.586) & (0.536) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 27094 & 32555 & 32828 & 24595 & 20567 & 26602 & 41575 & 34342 & 39826 & 38112 \\ 
turning point Dollar & 3357 & 4034 & 4068 & 3048 & 2549 & 3296 & 5152 & 4256 & 4935 & 4723 \\ 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 3,343 & 7,084 & 11,014 & 14,584 & 17,479 & 19,864 & 22,735 & 25,104 & 28,597 & 30,195 \\ 
R$^{2}$ & 0.454 & 0.386 & 0.374 & 0.357 & 0.353 & 0.354 & 0.355 & 0.352 & 0.349 & 0.345 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank of the city output share by SOE \
More specifically, the low deciles means a low SOE share by city \
Deciles close to one, however, implies a stronger share of SO2.
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}