### 2020-04-27 Temp_tables/Temp_tables_revision/05_Kuznet/03_kuznet_decile_capital.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Tue, Apr 07, 2020 - 12:09:59
\begin{table}[!htbp] \centering 
  \caption{Deciles capital} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-11} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9 &  Baseline\\
 \hline \\[-1.8ex] 
   \text{(ln gdp per cap)}_{ct}  & 4.446$^{**}$ & 2.936$^{*}$ & 2.483$^{**}$ & 2.819$^{**}$ & 3.089$^{***}$ & 3.076$^{***}$ & 2.876$^{***}$ & 2.456$^{***}$ & 2.573$^{***}$ & 2.363$^{***}$ \\ 
  & (2.070) & (1.684) & (1.210) & (1.119) & (0.990) & (0.886) & (0.851) & (0.791) & (0.766) & (0.752) \\ 
   \text{(ln gdp per cap) squared}_{ct}  & $-$0.218$^{**}$ & $-$0.141$^{*}$ & $-$0.119$^{**}$ & $-$0.140$^{**}$ & $-$0.156$^{***}$ & $-$0.151$^{***}$ & $-$0.135$^{***}$ & $-$0.118$^{***}$ & $-$0.121$^{***}$ & $-$0.112$^{***}$ \\ 
  & (0.098) & (0.081) & (0.060) & (0.055) & (0.048) & (0.043) & (0.041) & (0.038) & (0.037) & (0.037) \\ 
   \text{(ln population)}_{ct}  & 0.018 & 0.008 & 0.080 & 0.058 & $-$0.002 & 0.039 & 0.058 & 0.050 & 0.047 & 0.042 \\ 
  & (0.139) & (0.117) & (0.113) & (0.110) & (0.103) & (0.098) & (0.096) & (0.092) & (0.091) & (0.088) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$1.549$^{**}$ & $-$0.488 & $-$0.234 & $-$0.382$^{*}$ & $-$0.370$^{*}$ & $-$0.389$^{**}$ & $-$0.358$^{**}$ & $-$0.371$^{**}$ & $-$0.326$^{**}$ & $-$0.306$^{**}$ \\ 
  & (0.651) & (0.342) & (0.250) & (0.218) & (0.202) & (0.176) & (0.163) & (0.156) & (0.145) & (0.137) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  & 3.677 & 1.149 & 1.163 & 1.155 & 1.331$^{*}$ & 1.321$^{**}$ & 1.248$^{**}$ & 1.183$^{**}$ & 0.856$^{*}$ & 0.779$^{*}$ \\ 
  & (2.640) & (1.642) & (0.885) & (0.843) & (0.740) & (0.626) & (0.579) & (0.529) & (0.467) & (0.430) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 27158 & 32712 & 32813 & 24332 & 20435 & 26542 & 41615 & 34303 & 39839 & 38160 \\ 
turning point Dollar & 3365 & 4053 & 4066 & 3015 & 2532 & 3289 & 5157 & 4251 & 4937 & 4729 \\ 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 3,343 & 7,084 & 11,014 & 14,584 & 17,479 & 19,864 & 22,735 & 25,104 & 28,597 & 30,195 \\ 
R$^{2}$ & 0.453 & 0.386 & 0.373 & 0.357 & 0.353 & 0.354 & 0.355 & 0.352 & 0.349 & 0.345 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank of the city output share by SOE \
More specifically, the low deciles means a low SOE share by city \
Deciles close to one, however, implies a stronger share of SO2.
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}