### 2020-04-27 Temp_tables/Temp_tables_revision/05_Kuznet/04_kuznet_decile_emp.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Tue, Apr 07, 2020 - 12:10:21
\begin{table}[!htbp] \centering 
  \caption{Deciles employment} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-11} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& decile .1 & decile .2 &  decile .3 & decile .4 & decile .5 & decile .6 &  decile .7 & decile .8 & decile .9 &  Baseline\\
 \hline \\[-1.8ex] 
   \text{(ln gdp per cap)}_{ct}  & 4.413$^{**}$ & 2.961$^{*}$ & 2.487$^{**}$ & 2.818$^{**}$ & 3.093$^{***}$ & 3.087$^{***}$ & 2.885$^{***}$ & 2.477$^{***}$ & 2.589$^{***}$ & 2.383$^{***}$ \\ 
  & (2.071) & (1.685) & (1.211) & (1.120) & (0.989) & (0.884) & (0.849) & (0.788) & (0.763) & (0.750) \\ 
   \text{(ln gdp per cap) squared}_{ct}  & $-$0.216$^{**}$ & $-$0.142$^{*}$ & $-$0.120$^{**}$ & $-$0.139$^{**}$ & $-$0.156$^{***}$ & $-$0.151$^{***}$ & $-$0.136$^{***}$ & $-$0.119$^{***}$ & $-$0.122$^{***}$ & $-$0.113$^{***}$ \\ 
  & (0.098) & (0.081) & (0.060) & (0.055) & (0.048) & (0.043) & (0.041) & (0.038) & (0.037) & (0.037) \\ 
   \text{(ln population)}_{ct}  & 0.020 & 0.007 & 0.080 & 0.059 & $-$0.001 & 0.040 & 0.058 & 0.050 & 0.047 & 0.042 \\ 
  & (0.140) & (0.117) & (0.113) & (0.110) & (0.103) & (0.098) & (0.096) & (0.092) & (0.091) & (0.088) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$1.841$^{**}$ & $-$0.511 & $-$0.228 & $-$0.417$^{*}$ & $-$0.415$^{**}$ & $-$0.431$^{**}$ & $-$0.394$^{**}$ & $-$0.388$^{**}$ & $-$0.335$^{**}$ & $-$0.325$^{**}$ \\ 
  & (0.745) & (0.355) & (0.266) & (0.228) & (0.211) & (0.182) & (0.168) & (0.160) & (0.147) & (0.138) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  & 6.641 & 1.699 & 1.427 & 1.644 & 1.907$^{*}$ & 1.874$^{**}$ & 1.677$^{**}$ & 1.495$^{**}$ & 1.075$^{*}$ & 1.011$^{*}$ \\ 
  & (4.766) & (2.081) & (1.199) & (1.078) & (0.979) & (0.796) & (0.729) & (0.659) & (0.575) & (0.527) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 27513 & 32868 & 32912 & 24615 & 20598 & 26680 & 41534 & 34281 & 39743 & 38159 \\ 
turning point Dollar & 3409 & 4073 & 4078 & 3050 & 2552 & 3306 & 5147 & 4248 & 4925 & 4729 \\ 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 3,343 & 7,084 & 11,014 & 14,584 & 17,479 & 19,864 & 22,735 & 25,104 & 28,597 & 30,195 \\ 
R$^{2}$ & 0.453 & 0.386 & 0.373 & 0.357 & 0.353 & 0.354 & 0.355 & 0.352 & 0.349 & 0.345 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank of the city output share by SOE \
More specifically, the low deciles means a low SOE share by city \
Deciles close to one, however, implies a stronger share of SO2.
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}