### 2020-04-27 Temp_tables/Temp_tables_revision/05_Kuznet/05_kuznet_median.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Tue, Apr 07, 2020 - 19:25:45
\begin{table}[!htbp] \centering 
  \caption{Deciles employment} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-7} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\
 \\[-1.8ex]& Output & Output & Capital & Capital & Labour & Labour\\
 \hline \\[-1.8ex] 
   \text{(ln gdp per cap)}_{ct}  & 3.092$^{***}$ & 1.079 & 3.089$^{***}$ & 1.072 & 3.093$^{***}$ & 1.077 \\ 
  & (0.990) & (1.288) & (0.990) & (1.283) & (0.989) & (1.287) \\ 
   \text{(ln gdp per cap) squared}_{ct}  & $-$0.156$^{***}$ & $-$0.044 & $-$0.156$^{***}$ & $-$0.044 & $-$0.156$^{***}$ & $-$0.044 \\ 
  & (0.048) & (0.064) & (0.048) & (0.064) & (0.048) & (0.064) \\ 
   \text{(ln population)}_{ct}  & $-$0.00005 & 0.060 & $-$0.002 & 0.063 & $-$0.001 & 0.059 \\ 
  & (0.103) & (0.184) & (0.103) & (0.183) & (0.103) & (0.184) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.366$^{**}$ & $-$0.252 & $-$0.370$^{*}$ & $-$0.238 & $-$0.415$^{**}$ & $-$0.229 \\ 
  & (0.186) & (0.177) & (0.202) & (0.190) & (0.211) & (0.188) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 2.182$^{**}$ & 0.600 &  &  &  &  \\ 
  & (0.963) & (0.653) &  &  &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  &  & 1.331$^{*}$ & 0.364 &  &  \\ 
  &  &  & (0.740) & (0.525) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  &  &  & 1.907$^{*}$ & 0.356 \\ 
  &  &  &  &  & (0.979) & (0.637) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 20567 & 186475 & 20435 & 185728 & 20598 & 185170 \\ 
turning point Dollar & 2549 & 23107 & 2532 & 23015 & 2552 & 22946 \\ 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 17,479 & 12,716 & 17,479 & 12,716 & 17,479 & 12,716 \\ 
R$^{2}$ & 0.353 & 0.377 & 0.353 & 0.377 & 0.353 & 0.377 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
A decile indicates the rank of the city output share by SOE \
More specifically, the low deciles means a low SOE share by city \
Deciles close to one, however, implies a stronger share of SO2.
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}