### 2020-04-27 Temp_tables/Temp_tables_revision/06_table_8_city_industry/01_kuznet.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Wed, Apr 08, 2020 - 17:33:36
\begin{table}[!htbp] \centering 
  \caption{Soft budget constraint as reflected in the SOEs reaction to the  SPZ_c s and coastal area's policies in China } 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{3}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-4} 
\\[-1.8ex] & (1) & (2) & (3)\\
 \\[-1.8ex]& Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   \text{(ln gdp per cap)}_{ct}  & 2.365$^{***}$ & 2.363$^{***}$ & 2.383$^{***}$ \\ 
  & (0.751) & (0.752) & (0.750) \\ 
   \text{(ln gdp per cap) squared}_{ct}  & $-$0.112$^{***}$ & $-$0.112$^{***}$ & $-$0.113$^{***}$ \\ 
  & (0.037) & (0.037) & (0.037) \\ 
   \text{(ln population)}_{ct}  & 0.042 & 0.042 & 0.042 \\ 
  & (0.089) & (0.088) & (0.088) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & $-$0.303$^{**}$ & $-$0.306$^{**}$ & $-$0.325$^{**}$ \\ 
  & (0.127) & (0.137) & (0.138) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.160$^{**}$ &  &  \\ 
  & (0.536) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 0.779$^{*}$ &  \\ 
  &  & (0.430) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 1.011$^{*}$ \\ 
  &  &  & (0.527) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 38112 & 38160 & 38159 \\ 
turning point Dollar & 4723 & 4729 & 4729 \\ 
City fixed effects & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes \\ 
Yes fixed effects & Yes & Yes & Yes \\ 
Observations & 30,195 & 30,195 & 30,195 \\ 
R$^{2}$ & 0.345 & 0.345 & 0.345 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}