### 2020-04-27 Temp_tables/Temp_tables_revision/06_table_8_city_industry/02_table_8_model_A_panel_A.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Thu, Apr 09, 2020 - 16:34:00
\begin{table}[!htbp] \centering 
  \caption{Test Median city} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccccccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{15}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-16} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12) & (13) & (14) & (15)\\
 \\[-1.8ex]& Output & Capital & LabourOutput & Capital & Labour & Output & Capital & Labour & Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   TCZ_c \times \text{Period} \times \text{Polluted}_i  & 0.157 & 0.235 & 0.177 & 0.634 & 0.342 & 0.110 & 0.148 & 0.130 & 0.126 & $-$0.356 & $-$0.401 & $-$0.465 & $-$0.407$^{*}$ & $-$0.432$^{*}$ & $-$0.506$^{**}$ \\ 
  & (0.306) & (0.335) & (0.344) & (2.031) & (1.327) & (1.404) & (0.705) & (0.821) & (0.772) & (0.294) & (0.331) & (0.316) & (0.230) & (0.251) & (0.252) \\ 
   TCZ_c \times \text{Period} \times \text{output share SOE}_{i}  & $-$0.047 &  &  & 26.762 &  &  & 0.127 &  &  & $-$1.577$^{**}$ &  &  & $-$1.560 &  &  \\ 
  & (0.893) &  &  & (50.983) &  &  & (1.155) &  &  & (0.724) &  &  & (0.984) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & $-$0.808 &  &  & $-$25.251 &  &  & 0.214 &  &  & 2.199$^{**}$ &  &  & 2.485$^{**}$ &  &  \\ 
  & (1.391) &  &  & (49.787) &  &  & (1.587) &  &  & (0.878) &  &  & (1.198) &  &  \\ 
   TCZ_c \times \text{Period} \times \text{capital share SOE}_{i}  &  & 0.055 &  &  & 11.201 &  &  & $-$0.048 &  &  & $-$1.432$^{**}$ &  &  & $-$0.945 &  \\ 
  &  & (0.745) &  &  & (10.042) &  &  & (1.156) &  &  & (0.659) &  &  & (0.637) &  \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & $-$0.901 &  &  & $-$9.770 &  &  & 0.213 &  &  & 1.844$^{**}$ &  &  & 1.635$^{*}$ &  \\ 
  &  & (1.089) &  &  & (10.175) &  &  & (1.582) &  &  & (0.920) &  &  & (0.837) &  \\ 
   TCZ_c \times \text{Period} \times \text{labour share SOE}_{i}  &  &  & $-$0.079 &  &  & 9.328 &  &  & 0.270 &  &  & $-$1.721$^{**}$ &  &  & $-$1.556$^{*}$ \\ 
  &  &  & (0.895) &  &  & (18.116) &  &  & (1.281) &  &  & (0.854) &  &  & (0.875) \\ 
   TCZ_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & $-$0.721 &  &  & $-$7.955 &  &  & 0.223 &  &  & 2.416$^{**}$ &  &  & 2.489$^{**}$ \\ 
  &  &  & (1.349) &  &  & (18.130) &  &  & (1.612) &  &  & (0.992) &  &  & (1.084) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Yes fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 9,733 & 9,733 & 9,733 & 2,346 & 2,346 & 2,346 & 4,022 & 4,022 & 4,022 & 14,513 & 14,513 & 14,513 & 17,915 & 17,915 & 17,915 \\ 
R$^{2}$ & 0.900 & 0.900 & 0.900 & 0.951 & 0.951 & 0.951 & 0.936 & 0.936 & 0.936 & 0.846 & 0.846 & 0.846 & 0.852 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}