### 2020-04-27 Temp_tables/Temp_tables_revision/06_table_8_city_industry/04_table_8_model_B_panel_A.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{booktabs,caption,threeparttable, siunitx, adjustbox}

\begin{document}% Date and time \times Thu, Apr 09, 2020 - 16:16:49
\begin{table}[!htbp] \centering 
  \caption{Test Median city} 
\label{}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccccccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{15}{c}{Dependent variable \times \text { SO2 emission }_{i k t}} \\ 
\cline{2-16} 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12) & (13) & (14) & (15)\\
 \\[-1.8ex]& Output & Capital & Labour & Output & Capital & Labour & Output & Capital & Labour & Output & Capital & Labour & Output & Capital & Labour\\
 \hline \\[-1.8ex] 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.559 & $-$0.590 & $-$0.665 & 0.624 & 2.272 & 1.013 & $-$0.927 & $-$1.262 & $-$1.072 & $-$0.439 & $-$0.439 & $-$0.481 & $-$0.410 & $-$0.436 & $-$0.474 \\ 
  & (0.691) & (0.741) & (0.749) & (4.839) & (5.004) & (5.344) & (1.755) & (1.885) & (1.991) & (0.275) & (0.304) & (0.307) & (0.305) & (0.335) & (0.343) \\ 
   target_c \times \text{Period} \times \text{output share SOE}_{i}  & $-$1.076 &  &  & 20.489 &  &  & 0.079 &  &  & 0.006 &  &  & $-$1.040 &  &  \\ 
  & (2.046) &  &  & (20.662) &  &  & (2.809) &  &  & (0.863) &  &  & (0.902) &  &  \\ 
   target_c \times \text{Period} \times \text{Polluted}_i \times \text{output share SOE}_{i}  & 1.665 &  &  & $-$28.886 &  &  & 0.845 &  &  & 0.161 &  &  & 0.779 &  &  \\ 
  & (2.434) &  &  & (35.232) &  &  & (3.053) &  &  & (1.077) &  &  & (1.250) &  &  \\ 
   target_c \times \text{Period} \times \text{capital share SOE}_{i}  &  & $-$0.732 &  &  & 10.478 &  &  & $-$0.558 &  &  & 0.062 &  &  & $-$0.654 &  \\ 
  &  & (1.613) &  &  & (11.104) &  &  & (2.534) &  &  & (0.676) &  &  & (0.638) &  \\ 
   target_c \times \text{Period} \times \text{Polluted}_i \times \text{capital share SOE}_{i}  &  & 1.273 &  &  & $-$27.452 &  &  & 1.423 &  &  & 0.097 &  &  & 0.578 &  \\ 
  &  & (2.025) &  &  & (25.674) &  &  & (2.789) &  &  & (0.931) &  &  & (1.000) &  \\ 
   target_c \times \text{Period} \times \text{labour share SOE}_{i}  &  &  & $-$1.066 &  &  & 13.662 &  &  & 0.153 &  &  & $-$0.016 &  &  & $-$0.992 \\ 
  &  &  & (2.191) &  &  & (14.363) &  &  & (3.200) &  &  & (0.796) &  &  & (0.810) \\ 
   target_c \times \text{Period} \times \text{Polluted}_i \times \text{labour share SOE}_{i}  &  &  & 1.836 &  &  & $-$22.235 &  &  & 1.047 &  &  & 0.332 &  &  & 0.989 \\ 
  &  &  & (2.533) &  &  & (29.210) &  &  & (3.438) &  &  & (1.033) &  &  & (1.203) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Yes fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 6,496 & 6,496 & 6,496 & 2,346 & 2,346 & 2,346 & 4,022 & 4,022 & 4,022 & 14,513 & 14,513 & 14,513 & 17,915 & 17,915 & 17,915 \\ 
R$^{2}$ & 0.925 & 0.925 & 0.925 & 0.951 & 0.952 & 0.951 & 0.937 & 0.937 & 0.937 & 0.846 & 0.846 & 0.846 & 0.852 & 0.852 & 0.852 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1-3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\end{document}