### 2020-05-09 Paper/table_.tex

\begin{tabular}{lrrrrrrrr}
\toprule
{} & \multicolumn{2}{l}{difference} & \multicolumn{2}{l}{variance} & \multicolumn{2}{l}{difference} & \multicolumn{2}{l}{variance} \\
TCZ &     No TCZ &     TCZ &   No TCZ &  TCZ & No SOE dominated & SOE dominated & No SOE dominated & SOE dominated \\
Location    &            &         &          &      &                  &               &                  &               \\
\midrule
Full sample &    -192 & -436 &     0.22 & 0.26 &          -350 &       -327 &             0.24 &          0.28 \\
Central     &    -238 & -232 &     0.25 & 0.19 &          -250 &       -214 &             0.24 &          0.18 \\
Coastal     &    -138 & -659 &     0.12 & 0.29 &          -549 &       -344 &             0.27 &          0.25 \\
Northeast   &    -206 &  -89 &     0.30 & 0.08 &          -159 &       -163 &             0.19 &          0.19 \\
Northwest   &    -115 & -277 &     0.21 & 0.31 &            -2 &       -312 &             0 &          0.39 \\
Southwest   &    -272 & -503 &     0.31 & 0.29 &          -247 &       -641 &             0.18 &          0.38 \\
Non Coastal &    -215 & -302 &     0.28 & 0.23 &          -198 &       -331 &             0.20 &          0.29 \\
Coastal     &    -140 & -578 &     0.13 & 0.28 &          -484 &       -314 &             0.26 &          0.23 \\
\bottomrule
\end{tabular}

	         No_TCZ	  TCZ	            No SOE dominated               dominated	     
tso2_cit	2,691,064	4,923,903	4,358,489	3,570,123

TCZ
https://quicklatex.com/cache3/eb/ql_135e8f554e9af28e18d27309d086bbeb_l3.png

SOE
https://quicklatex.com/cache3/20/ql_ac6c39112b2a3223557aad5fdb83b220_l3.png

2006: 2588 
2007:

TCZ_c	No_TCZ	TCZ	difference_no_tcz	difference_tcz
Period	After	Before	After	Before		
tso2_cit	586975721	485819107	1728326765	1446874686	101156614	281452079

1222604206

https://www.google.com/search?newwindow=1&safe=active&sxsrf=ALeKk02a5_rldLmU_FSo5jcjmBPbk_Toog%3A1588334031328&ei=zw2sXsTBE4uLlwSpzaP4Cg&q=kilos+to+tonnes&oq=kilos+to+tonn&gs_lcp=CgZwc3ktYWIQAxgBMgcIIxCxAhAnMgQIABBDMgIIADIHCAAQFBCHAjICCAAyBAgAEAoyBAgAEEMyBAgAEAoyBAgAEEMyBAgAEAo6BAgAEEc6BAgjECdQ8wJYlgdgwQ5oAHACeACAAZECiAHVA5IBBTAuMS4xmAEAoAEBqgEHZ3dzLXdpeg&sclient=psy-ab