### 2020-05-09 Temp_tables/Tables_paper/02_paper_version_2/00_all_tables.tex

\documentclass[12pt]{article} 
\usepackage[utf8]{inputenc}
\usepackage{amssymb,amsmath,amsfonts,eurosym,geometry,
ulem,graphicx,color,setspace,sectsty,comment,footmisc,
natbib,
pdflscape,subfigure,array,hyperref, booktabs,
threeparttable, siunitx, adjustbox,
%rotating
}

\begin{document}

\begin{table}[!htbp] \centering
\caption{\\ SO2 reduction during the subsequent FYPs}
\label{table_1}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{lrrrr}
\toprule
             Cities &  1998-2001 &  2002-2005 &  2006-2010 &  Target \\
\midrule
             No TCZ &     21.00\% &     64.00\% &    -11.00\% &  -6.00\% \\
                TCZ &     -7.00\% &     38.00\% &    -15.00\% & -16.00\% \\
 No Dominated SOE^ a &    -31.00\% &     30.00\% &    -16.00\% & -10.00\% \\
    Dominated SOE^ a &      6.00\% &     21.00\% &    -17.00\% & -12.00\% \\
        Full Sample &     -2.00\% &     45.00\% &    -13.00\% & -10.00\% \\
\bottomrule
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\

Sources: Author's own computation \

The list of TCZ is provided by the State Council, 1998.
"Official Reply to the State Council Concerning Acid Rain Control Areas
and Sulfur Dioxide Pollution Control Areas".
The information about the SO2 level are collected using various edition
of the China Environment Statistics Yearbook.
We compute the reduction of SO2 emission using the same methodology
as Chen and al.(2018). \

$a$ (No) Dominated SOEs cities refer to cities where the 
(output, capital, employment) share of SOEs is (below) above a critical threshold,
for instance the 60th decile

\end{tablenotes}
\end{table}

\begin{equation} \label{eq:equation_1}
\begin{aligned} 
\text {Log SO2 emission }_{i k t}=& \alpha (Target_{i} \times \text { Polluted sectors }_{k} \times \text { Period }) 
& + \\ \theta {X}_{i k t}+\nu_{ik}+\lambda_{it} +\phi_{kt} +\epsilon_{ikt}
\end{aligned}
\end{equation}

\begin{table}[!htbp] \centering
\caption{\\ Summary statistics by city characteristics}
\label{table_2}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{lrrr}
\toprule
index & Output share SOE_i & Capital share SOE_i & Employment share SOE_i \\
\midrule
Full sample          &             24.95\% &              35.48\% &                 29.19\% \\
Central              &             28.78\% &              41.80\% &                 34.17\% \\
Coastal              &             11.28\% &              18.50\% &                 12.04\% \\
Northeast            &             26.16\% &              38.73\% &                 33.83\% \\
Northwest            &             43.38\% &              51.06\% &                 49.53\% \\
Southwest            &             31.02\% &              44.59\% &                 36.08\% \\
Central              &             28.78\% &              41.80\% &                 34.17\% \\
Eastern              &             15.99\% &              24.83\% &                 19.05\% \\
Western              &             37.71\% &              48.44\% &                 42.82\% \\
No TCZ               &             24.32\% &              35.07\% &                 29.97\% \\
TCZ                  &             25.34\% &              35.74\% &                 28.70\% \\
Concentrated city    &             34.09\% &              45.39\% &                 37.89\% \\
No Concentrated city &             15.14\% &              24.86\% &                 19.86\% \\
\bottomrule
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\

Sources: Author's own computation

The list of TCZ is provided by the State Council, 1998.

Output $\text { Share SOE }_{i}$ refers to the ratio of output
(respectively capital, employment) of SOEs over the total production
(capital, employment) in city $i$

\end{tablenotes}
\end{table}

\begin{table}[!htbp] \centering
    \caption{\\ Summary statistics by city characteristics}
      \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
    \label{tab:table_3}
    \begin{tabular}{lrr}
      \multicolumn{1}{l}{\textbf{\small Panel A: Panel A: Pollution emission and Macro-variables}} \\
      \toprule
      & \multicolumn{1}{c}{No TCZ} & \multicolumn{1}{c}{TCZ} \\
      \midrule
      $SO2_{ikt}$  & 182,873 & 172,032 \\
      $\text{SO2 over population}_{ikt}$ &   2,754 &   1,633 \\
      $\text{gdp per capita}_i$ &  19,796 &  32,796 \\
      $\text{population}_i$ &      99 &     266 \\
      \bottomrule
      \\ %%% Create second table
      \multicolumn{1}{l}{\textbf{\small Panel B: SO2 reduction over 2000-2010}} \\
      \toprule
      {} & \multicolumn{2}{l}{\footnotesize difference (10.000 tons units)} \\
                  & No TCZ   & TCZ     \\
      Location    &          &         \\
      \midrule
      Full sample &    -192 & -436  \\
      Central     &    -238 & -232 \\
      Coastal     &    -138 & -659  \\
      Northeast   &    -206 &  -89 \\
      Northwest   &    -115 & -277  \\
      Southwest   &    -272 & -503 \\
      Non Coastal &    -215 & -302  \\
      Coastal     &    -140 & -578  \\
      \bottomrule
      \hline
    \end{tabular}
    \end{adjustbox}
    \begin{tablenotes}
      \small
      \item 
      Sources: Author's own computation \\
      Panel A provides summary statistics for the main variables used in the subsequent empirical analysis. \\
  Panel B provides summary statistics for the SO2 emission (in 10.000 tons) for Chinese cities (228), split into TCZ (140) and non-TCZ (88) cities \\
  $SO2_{ikt}$ is in kilos and $\text{population}_i$ in million
      \\
    \end{tablenotes}
\end{table}

\begin{equation} \label{eq:equation_2}
\Delta SO2_{i, 05 − 10}=\Delta SO2_{p, 05 − 10} \times \sum_{k=1}^{29} \mu_{k} \frac{\text { output value of industry } k \text { in city } i}{\text { output value of industry } k \text { in province } p}
\end{equation}

\begin{table}[!htbp] \centering
\caption{Summary statistics of Target by city characteristics}
\label{table_4}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{lrrr}
\toprule
{} & All Cities & No SOE dominated & SOE dominated \\
index                &            &                  &               \\
\midrule
Full sample          &     12.07\% &           11.89\% &        12.39\% \\
Central              &      8.97\% &            7.47\% &        10.99\% \\
Coastal              &     17.79\% &           18.04\% &        16.05\% \\
Northeast            &      5.76\% &            5.17\% &         6.81\% \\
Northwest            &      5.70\% &            2.87\% &         7.52\% \\
Southwest            &     16.83\% &           11.82\% &        21.53\% \\
Central              &      8.97\% &            7.47\% &        10.99\% \\
Eastern              &     14.13\% &           14.73\% &        11.52\% \\
Western              &     11.93\% &            8.61\% &        14.20\% \\
No TCZ               &      5.69\% &            5.03\% &         7.20\% \\
TCZ                  &     16.07\% &           16.75\% &        14.99\% \\
Concentrated city    &      9.72\% &            6.11\% &        12.87\% \\
No Concentrated city &     14.58\% &           15.34\% &        10.72\% \\
Coastal              &     17.79\% &           18.04\% &        16.05\% \\
No Coastal           &      9.64\% &            7.60\% &        11.83\% \\
\bottomrule
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\

Sources: Author's own computation

The list of TCZ is provided by the State Council, 1998. 
(No) SOE Dominated cities refers to cities where the 
(output, capital, employment) share of SOEs is (below) above a critical threshold,
for instance the 60th decile

(No) Concentrated city refers to cities where the 
Herfindahl index is (below) above a critical threshold,
for instance the 60th decile

\end{tablenotes}
\end{table}

\begin{table}[!htbp] \centering
  %\resizebox{1\textwidth}{!}{
    %\begin{threeparttable}
    \caption{Baseline Regression: State versus Private sectors}
      \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
     \label{table_5}
      \begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
        \multicolumn{1}{l}{\textbf{Panel A: SOE Dominating sectors}} \\
        \toprule
        & \multicolumn{8}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
        \cline{2-9}

\\[-1.8ex]
            &\multicolumn{2}{c}{Full sample}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{Employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
  output_{cit} & 0.144 & $-$0.066 & 0.875$^{*}$ & 0.216 & 1.092$^{**}$ & 0.308 & 0.995$^{**}$ & 0.087 \\ 
  & (0.152) & (0.089) & (0.457) & (0.257) & (0.464) & (0.397) & (0.421) & (0.280) \\ 
  capital_{cit} & 1.481$^{**}$ & 0.971$^{***}$ & $-$3.614$^{***}$ & $-$0.611 & $-$4.106$^{***}$ & $-$0.485 & $-$4.094$^{***}$ & $-$0.393 \\ 
  & (0.730) & (0.366) & (1.088) & (0.617) & (1.146) & (0.773) & (1.104) & (0.708) \\ 
  labour_{cit} & 3.089$^{***}$ & 1.538$^{**}$ & 11.671$^{***}$ & 4.601$^{***}$ & 12.228$^{***}$ & 5.216$^{***}$ & 12.327$^{***}$ & 4.783$^{***}$ \\ 
  & (0.832) & (0.767) & (2.326) & (1.475) & (2.374) & (1.353) & (2.359) & (1.548) \\ 
   target_c \times \text{Period}  & $-$0.003 &  & 0.087 &  & 0.012 &  & 0.128 &  \\ 
  & (0.104) &   & (0.381) &   & (0.379) &   & (0.379) &   \\ 
   target_c \times \text{Polluted}_i  & 0.438$^{***}$ &  & 0.838$^{**}$ &  & 0.748$^{*}$ &  & 0.839$^{**}$ &  \\ 
  & (0.140) &   & (0.383) &   & (0.390) &   & (0.379) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.291$^{**}$ & $-$0.430$^{***}$ & $-$0.430 & 0.009 & $-$0.344 & $-$0.159 & $-$0.468 & $-$0.215 \\ 
  & (0.146) & (0.132) & (0.465) & (0.465) & (0.443) & (0.435) & (0.432) & (0.461) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Observations & 30,676 & 30,676 & 9,165 & 9,165 & 9,149 & 9,149 & 9,011 & 9,011 \\ 
R$^{2}$ & 0.346 & 0.851 & 0.377 & 0.872 & 0.376 & 0.868 & 0.373 & 0.869 \\ 
        \bottomrule
        \\ %%% Create second table
        \multicolumn{1}{l}{\textbf{Panel B: Prviate Dominating sectors}} \\
        \toprule
        & \multicolumn{8}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
        \cline{2-9}

\\[-1.8ex]
            &\multicolumn{2}{c}{Full sample}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{Employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
  output_{cit} & 0.144 & $-$0.066 & 0.116 & $-$0.077 & 0.107 & $-$0.077 & 0.113 & $-$0.073 \\ 
  & (0.152) & (0.089) & (0.134) & (0.094) & (0.130) & (0.090) & (0.130) & (0.095) \\ 
  capital_{cit} & 1.481$^{**}$ & 0.971$^{***}$ & 1.885$^{***}$ & 0.942$^{**}$ & 1.930$^{***}$ & 0.905$^{**}$ & 1.744$^{***}$ & 0.911$^{**}$ \\ 
  & (0.730) & (0.366) & (0.575) & (0.445) & (0.557) & (0.401) & (0.551) & (0.441) \\ 
  labour_{cit} & 3.089$^{***}$ & 1.538$^{**}$ & 2.592$^{***}$ & 1.423$^{*}$ & 2.567$^{***}$ & 1.329$^{*}$ & 2.611$^{***}$ & 1.376 \\ 
  & (0.832) & (0.767) & (0.718) & (0.798) & (0.707) & (0.778) & (0.731) & (0.847) \\ 
   target_c \times \text{Period}  & $-$0.003 &  & $-$0.050 &  & $-$0.031 &  & $-$0.049 &  \\ 
  & (0.104) &   & (0.105) &   & (0.105) &   & (0.105) &   \\ 
   target_c \times \text{Polluted}_i  & 0.438$^{***}$ &  & 0.388$^{***}$ &  & 0.394$^{***}$ &  & 0.369$^{**}$ &  \\ 
  & (0.140) &   & (0.148) &   & (0.145) &   & (0.150) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.291$^{**}$ & $-$0.430$^{***}$ & $-$0.250$^{*}$ & $-$0.429$^{***}$ & $-$0.284$^{*}$ & $-$0.430$^{***}$ & $-$0.256$^{*}$ & $-$0.434$^{***}$ \\ 
  & (0.146) & (0.132) & (0.150) & (0.141) & (0.151) & (0.140) & (0.151) & (0.144) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Observations & 30,676 & 30,676 & 21,511 & 21,511 & 21,527 & 21,527 & 21,665 & 21,665 \\ 
R$^{2}$ & 0.346 & 0.851 & 0.355 & 0.855 & 0.359 & 0.856 & 0.358 & 0.856 \\ 
    \end{tabular}
    \end{adjustbox}
    \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedastiikty-robust standard errors in parentheses are clustered by city \\
      \\
    \end{tablenotes}
\end{table}

\begin{equation} \label{eq:equation_3}
 \text {Log SO2 emission }_{i k t}=\sum_{t=2002}^{2007} \alpha (Target_{i}  \times \text { Polluted sectors }_{k} \times year _{t}) +\theta {X}_{i k t}+\nu_{ik}+\lambda_{it} +\phi_{kt} +\epsilon_{ikt} 
\end{equation}

\begin{table}[!htbp] \centering 
  \caption{Parallel trend} 
\label{table_6}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{7}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-8}

\\[-1.8ex]
            &\multicolumn{1}{c}{}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7)\\
 \\[-1.8ex]& Full sample & No SOE & SOE & No SOE & SOE & No SOE & SOE\\
 \hline \\[-1.8ex] 
  2003 & $-$0.230 & $-$0.232 & $-$0.581 & $-$0.214 & $-$0.326 & $-$0.250 & $-$0.308 \\ 
  & (0.222) & (0.248) & (0.838) & (0.243) & (0.833) & (0.242) & (0.846) \\ 
  2004 & $-$0.156 & $-$0.174 & 0.148 & $-$0.197 & 0.498 & $-$0.229 & 0.498 \\ 
  & (0.222) & (0.240) & (0.808) & (0.237) & (0.772) & (0.239) & (0.807) \\ 
  2005 & $-$0.347 & $-$0.364 & $-$0.290 & $-$0.350 & $-$0.137 & $-$0.307 & $-$0.336 \\ 
  & (0.258) & (0.277) & (0.886) & (0.273) & (0.884) & (0.271) & (0.905) \\ 
  2006 & $-$0.598$^{**}$ & $-$0.645$^{**}$ & $-$0.082 & $-$0.625$^{**}$ & $-$0.019 & $-$0.630$^{**}$ & $-$0.246 \\ 
  & (0.263) & (0.279) & (0.950) & (0.276) & (0.956) & (0.276) & (0.978) \\ 
  2007 & $-$0.715$^{***}$ & $-$0.686$^{**}$ & $-$0.308 & $-$0.703$^{**}$ & $-$0.247 & $-$0.712$^{**}$ & $-$0.273 \\ 
  & (0.265) & (0.280) & (0.926) & (0.276) & (0.902) & (0.282) & (0.947) \\ 
 \hline \\[-1.8ex] 
City-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry-year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
City-industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 30,676 & 21,511 & 9,165 & 21,527 & 9,149 & 21,665 & 9,011 \\ 
R$^{2}$ & 0.852 & 0.855 & 0.872 & 0.856 & 0.868 & 0.856 & 0.869 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\begin{table}[!htbp] \centering 
  \caption{\\ Effects of the policies: TCZ, Go West, and SPZ } 
\label{table_7}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{12}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-13}

\\[-1.8ex]
            &\multicolumn{2}{c}{TCZ}&\multicolumn{2}{c}{No TCZ}&\multicolumn{2}{c}{ Coastal }&\multicolumn{2}{c}{No  Coastal}&\multicolumn{2}{c}{ SPZ }&\multicolumn{2}{c}{No  SPZ}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10) & (11) & (12)\\ 
\hline \\[-1.8ex] 
   target_c \times \text{Period}  & $-$0.012 &  & 0.899 &  & $-$0.202$^{*}$ &  & 0.470$^{**}$ &  & $-$0.031 &  & 0.731$^{***}$ &  \\ 
  & (0.111) &   & (0.874) &   & (0.111) &   & (0.224) &   & (0.114) &   & (0.225) &   \\ 
   target_c \times \text{Polluted}_i  & 0.388$^{***}$ &  & 2.570$^{**}$ &  & 0.321$^{**}$ &  & 0.661$^{***}$ &  & 0.446$^{***}$ &  & 0.688$^{**}$ &  \\ 
  & (0.146) &   & (1.182) &   & (0.161) &   & (0.238) &   & (0.158) &   & (0.272) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.254 & $-$0.413$^{***}$ & $-$1.335 & $-$1.415 & $-$0.241 & $-$0.356$^{**}$ & $-$0.424 & $-$0.714$^{**}$ & $-$0.317$^{*}$ & $-$0.412$^{***}$ & $-$0.695$^{**}$ & $-$0.844$^{***}$ \\ 
  & (0.155) & (0.138) & (1.128) & (1.059) & (0.174) & (0.157) & (0.274) & (0.278) & (0.177) & (0.157) & (0.302) & (0.324) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes & No & Yes \\ 
Observations & 23,333 & 23,333 & 7,343 & 7,343 & 17,915 & 17,915 & 12,761 & 12,761 & 14,513 & 14,513 & 16,163 & 16,163 \\ 
R$^{2}$ & 0.346 & 0.847 & 0.416 & 0.892 & 0.359 & 0.852 & 0.373 & 0.870 & 0.340 & 0.846 & 0.377 & 0.871 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1, 3, 5, 7, 9, 11).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\begin{table}[!htbp] \centering 
  \caption{\\ Industrial concentration and environmental regulation effectiveness} 
\label{table_8}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-5}

\\[-1.8ex]
            &\multicolumn{2}{c}{Concentrated}&\multicolumn{2}{c}{No Concentrated}\\
\\[-1.8ex] & (1) & (2) & (3) & (4)\\ 
\hline \\[-1.8ex] 
   target_c \times \text{Period}  & $-$0.156 &  & 0.012 &  \\ 
  & (0.323) &   & (0.104) &   \\ 
   target_c \times \text{Polluted}_i  & 0.532 &  & 0.411$^{***}$ &  \\ 
  & (0.351) &   & (0.145) &   \\ 
   target_c \times \text{Period} \times \text{Polluted}_i  & $-$0.074 & 0.208 & $-$0.302$^{**}$ & $-$0.504$^{***}$ \\ 
  & (0.392) & (0.396) & (0.153) & (0.143) \\ 
 \hline \\[-1.8ex] 
City fixed effects & Yes & No & Yes & No \\ 
Industry fixed effects & Yes & No & Yes & No \\ 
Year fixed effects & Yes & No & Yes & No \\ 
City-year fixed effects & No & Yes & No & Yes \\ 
Industry-year fixed effects & No & Yes & No & Yes \\ 
City-industry fixed effects & No & Yes & No & Yes \\ 
Observations & 9,181 & 9,181 & 21,495 & 21,495 \\ 
R$^{2}$ & 0.406 & 0.888 & 0.342 & 0.846 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
for the regressions with city,industry, year fixed effect (i.e. columns 1 and 3).
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}

\begin{equation} \label{eq:equation_4}
 \text {Log SO2 emission }_{i k t} = \alpha \text{(ln gdp per cap)}_{ct}  + \\
 \beta\text{(ln gdp per cap)}^2_{ct}  + \gamma \text{(ln population)}_{ct}+\nu_{c}+\lambda_{k}+\phi_{t}+\epsilon_{i k t}
\end{equation}

\begin{sidewaystable}%[!htb] \centering
\begin{table}[!htbp] \centering 
  \caption{\\ Turning points and Kuznets curves} 
\label{table_9}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lcccccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{10}{c}{Dependent variable \text { SO2 emission }_{i k t}} \\ 
\cline{2-11}

\\[-1.8ex]
            &\multicolumn{2}{c}{City}&\multicolumn{2}{c}{Concentration}&\multicolumn{2}{c}{Output}&\multicolumn{2}{c}{Capital}&\multicolumn{2}{c}{Employment}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8) & (9) & (10)\\
 \\[-1.8ex]& TCZ & No TCZ & Concentrated & No Concentrated & SOE dominated & SOE No dominated & SOE dominated & SOE No dominated & SOE dominated & SOE No dominated\\
 \hline \\[-1.8ex] 
  \text{(ln gdp per cap)}_{ct}  & 2.708$^{***}$ & 0.269 & $-$0.626 & 3.391$^{***}$ & $-$0.096 & 2.611$^{***}$ & 1.437 & 2.559$^{***}$ & 0.434 & 2.984$^{***}$ \\ 
  & (0.927) & (1.428) & (1.192) & (0.971) & (1.784) & (0.837) & (1.882) & (0.847) & (1.555) & (0.854) \\ 
   \text{(ln gdp per cap)}^2_{ct}  & $-$0.132$^{***}$ & 0.002 & 0.039 & $-$0.158$^{***}$ & 0.017 & $-$0.127$^{***}$ & $-$0.062 & $-$0.126$^{***}$ & $-$0.015 & $-$0.143$^{***}$ \\ 
  & (0.045) & (0.071) & (0.061) & (0.047) & (0.090) & (0.041) & (0.096) & (0.041) & (0.080) & (0.042) \\ 
   \text{(ln population)}_{ct}  & 0.062 & 0.011 & 0.328$^{*}$ & 0.026 & 0.043 & 0.040 & $-$0.093 & 0.055 & 0.054 & 0.036 \\ 
  & (0.103) & (0.168) & (0.191) & (0.102) & (0.211) & (0.096) & (0.221) & (0.094) & (0.218) & (0.092) \\ 
 \hline \\[-1.8ex] 
turning point RMB & 28795 & - & - & 45396 & - & 30264 & - & 24867 & - & 35190 \\ 
turning point Dollar & 3568 & - & - & 5625 & - & 3750 & - & 3081 & - & 4361 \\ 
City fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Industry fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Year fixed effects & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 22,865 & 7,330 & 9,153 & 21,042 & 9,160 & 21,035 & 9,137 & 21,058 & 8,993 & 21,202 \\ 
R$^{2}$ & 0.344 & 0.415 & 0.405 & 0.340 & 0.376 & 0.353 & 0.375 & 0.357 & 0.372 & 0.357 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ 
\footnotesize{
Due to limited space, only the coefficients of interest are presented 
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}

\end{tablenotes}
\end{table}
\end{sidewaystable}

\begin{equation} \label{eq:equation_5}
TFP_{fikt}=\alpha\left(\text { Target }_{i} \times \text { Polluted sectors }_{k} \times \text { Period }\right)+ \zeta_{f}+ \nu_{i k}+\lambda_{i t}+\phi_{k t}+\epsilon_{i k t}
\end{equation}

\begin{table}[!htbp] \centering
  %\resizebox{1\textwidth}{!}{
    %\begin{threeparttable}
    \caption{Reduction mandate and TFP}
      \begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
     \label{}
      \begin{tabular}{@{\extracolsep{5pt}}lcccc}  
        \multicolumn{1}{l}{\textbf{Panel A: TCZ versus non-TCZ}} \\
        \toprule
        & \multicolumn{4}{c}{Dependent variable \text { TFP }_{fikt}} \\ 
\cline{2-5}

\\[-1.8ex]
            &\multicolumn{2}{c}{SOE}&\multicolumn{2}{c}{PRIVATE}\\
\\[-1.8ex] & (1) & (2) & (3) & (4)\\
 \\[-1.8ex]& TCZ & NO TCZ & TCZ & NO TCZ\\
 \hline \\[-1.8ex] 
   target_c \times \text{Period} \times \text{Polluted}_i  & 0.144$^{***}$ & $-$0.419 & $-$0.022 & $-$0.421$^{**}$ \\ 
  & (0.050) & (0.429) & (0.021) & (0.188) \\ 
 \hline \\[-1.8ex] 
Firm & Yes & Yes & Yes & Yes \\ 
City-industry &Yes & Yes & Yes & Yes \\ 
City-time & Yes & Yes & Yes & Yes \\ 
time-industry & Yes & Yes & Yes & Yes \\ 
Observations & 32,078 & 9,410 & 517,652 & 89,657 \\ 
R$^{2}$ & 0.953 & 0.961 & 0.861 & 0.869 \\

        \bottomrule
        \\ %%% Create second table
        \multicolumn{1}{l}{\textbf{Panel B: Coastal  versus non - Coastal}} \\
        \toprule
        & \multicolumn{4}{c}{Dependent variable \text { TFP }_{fikt}} \\ 
\cline{2-5}

\\[-1.8ex]
            &\multicolumn{2}{c}{SOE}&\multicolumn{2}{c}{PRIVATE}\\
\\[-1.8ex] & (1) & (2) & (3) & (4)\\
 \\[-1.8ex]&  Coastal_c  & NO  Coastal_c  &  Coastal_c  & NO  Coastal_c \\
 \hline \\[-1.8ex] 
   target_c \times \text{Period} \times \text{Polluted}_i  & 0.158$^{**}$ & 0.119 & $-$0.012 & $-$0.087$^{**}$ \\ 
  & (0.063) & (0.098) & (0.023) & (0.036) \\ 
 \hline \\[-1.8ex] 
Firm & Yes & Yes & Yes & Yes \\ 
City-industry & Yes & Yes & Yes & Yes \\ 
City-time & Yes & Yes & Yes & Yes \\ 
time-industry & Yes & Yes & Yes & Yes \\ 
Observations & 19,540 & 21,948 & 477,084 & 130,225 \\ 
R$^{2}$ & 0.955 & 0.956 & 0.857 & 0.878 \\

\bottomrule 
\\ %%% Create second table
        \multicolumn{1}{l}{\textbf{Panel C: industrial concentration}} \\
        \toprule
         & \multicolumn{4}{c}{Dependent variable \text { TFP }_{fikt}} \\ 
\cline{2-5}

\\[-1.8ex]
            &\multicolumn{2}{c}{SOE}&\multicolumn{2}{c}{PRIVATE}\\
\\[-1.8ex] & (1) & (2) & (3) & (4)\\
 \\[-1.8ex]& Concentrated & NO Concentrated & Concentrated & NO Concentrated\\
 \hline \\[-1.8ex] 
   target_c \times \text{Period} \times \text{Polluted}_i  & 0.068 & 0.159$^{**}$ & $-$0.035 & $-$0.015 \\ 
  & (0.084) & (0.063) & (0.032) & (0.024) \\ 
 \hline \\[-1.8ex] 
Firm & Yes & Yes & Yes & Yes \\ 
City-industry & Yes & Yes & Yes & Yes \\ 
City-time & Yes & Yes & Yes & Yes \\ 
time-industry & Yes & Yes & Yes & Yes \\ 
Observations & 23,054 & 18,434 & 170,305 & 437,004 \\ 
R$^{2}$ & 0.957 & 0.953 & 0.869 & 0.859 \\ 
    \end{tabular}
    \end{adjustbox}
    \begin{tablenotes}
      \small
      \item 
      Note: $^{*}$p$<$0.1 $^{**}$p$<$0.05 $^{***}$p$<$0.01 \\
      Heteroskedasticity-robust standard errors in parentheses are clustered by city \\
      \\
    \end{tablenotes}
\end{table}

\begin{table}[!htbp] \centering 
  \caption{TFP: below and above turning points} 
\label{table_11}
\begin{adjustbox}{width=\textwidth, totalheight=\textheight-2\baselineskip,keepaspectratio}
\begin{tabular}{@{\extracolsep{5pt}}lccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{4}{c}{Dependent variable \text { TFP }_{fikt}} \\ 
\cline{2-6}

\\[-1.8ex]
            &\multicolumn{2}{c}{SOE}&\multicolumn{2}{c}{PRIVATE}\\
\\[-1.8ex] & (1) & (2) & (3) & (4) \\
 \\[-1.8ex]& Above & Below & Above & Below \\
 \hline \\[-1.8ex] 
target_c \times \text{Period} \times \text{Polluted}_i  &0.047 (0.102) & 0.024 (0.106) & -0.009 (0.020)& -0.008 (0.048) &  \\
Observations  & 5,130 & 35,622 & 186,699 & 413,836&  Column (4): No Concentrated^a\\
R$^{2}$  & 0.979&0.956&0.905&0.876&  RMB 45396\\
\hline 
target_c \times \text{Period} \times \text{Polluted}_i  &  0.117 (0.071) & 0.042 (0.116) & -0.018 (0.020) & -0.109 (0.069)&  \\
Observations  & 9.397&31,355 & 284,694 & 315,841 &  Column (10): SOE No dominated^a \\ 
R$^{2}$  &0.971&0.960&0.890&0.883&  RMB 35190\\
\hline 
target_c \times \text{Period} \times \text{Polluted}_i  &  0.168^{* * *}  (0.058) & 0.108 (0.126)  &-0.013 (0.020) & 0.124 (0.076)& \\
Observations  & 12,605 & 28,147 & 350,740 & 249,795 &  Column (6): SOE No dominated^a \\ 
R$^{2}$  & 0.965&0.963&0.882&0.892 &  RMB 30264\\
\hline 
target_c \times \text{Period} \times \text{Polluted}_i  & 
0.156$^{* * *}$ (0.056)& 0.178 (0.137)& $-$0.017  (0.021)& $-$0.146$^{*}$ (0.082) & \\
Observations  & 13,935 & 26,817 & 366,289 & 234,246 & Column (1): TCZ^a\\ 
R$^{2}$  &0.964&0.964&0.879&0.893&  RMB  28795\\ 
\hline 
target_c \times \text{Period} \times \text{Polluted}_i  & 0.123^{* *} (0.051) & 0.130 (0.143) & -0.017 (0.021)&  -0.161 (0.100)& \\
Observations  & 18,061&22,691 & 419,579 & 180,956&  Column (8): SOE No dominated^a \\ 
R$^{2}$  & 0.958&0.966&0.869&0.894&  RMB 24867\\
\hline \\[-1.8ex] 
\end{tabular}
\end{adjustbox}
\begin{tablenotes} 
 \small 
 \item \\ \footnotesize{
The columns Above (Below) refer to firms in cities whose GDP per Capita are strictly above (below) the Kuznets turning point. References are provided in the last column. \\
$a$ refers to the number of the column from table \ref{tab:table_9}, which provides us with the estimated turning point. For instance 45 396 is estimated using the sub-sample of firms in no-concentrated cities. \\
Due to limited space, only the coefficients of interest are presented 
\sym{*} Significance at the 10\%, \sym{**} Significance at the 5\%, \sym{***} Significance at the 1\% \\
heteroscedasticity-robust standard errors in parentheses are clustered by city 
}
\end{tablenotes}
\end{table}

\end{document}